---
title: "adult_wae_collation"
author: "Chris Rounds"
date: "10/29/2024"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(sf)
library(arrow)

select <- dplyr::select
options(scipen = 999)
set.seed(999)
sf_use_s2(FALSE) # we dont like spheres
```


# Read in some fun data
```{r}
aged <- open_dataset("./data/age_assigned_data/mn_halk_aged_data/part-0.parquet")
#glimpse(aged)

mn_lakes <- readRDS("./data/mndow_lakes_sf_allDataUntransformed.rds")

stocking <- read.csv("./data/mn_wae_stocking.csv") %>%
  rename(lake.id = DOW, year = Year) %>%
  mutate(non.fry = fgl.pa + adl.pa + yrl.pa + frl.pa,
         lake.id = mnsentinellakes::fixlakeid(lake.id),
         non.fry = ifelse(is.na(non.fry), 0, non.fry)) %>%
  dplyr::select(lake.id, year, non.fry)

stocking_summarized <- stocking %>% group_by(lake.id, year) %>%
  summarize(non.fry = sum(non.fry), latest = max(year))

```


# filter for appropriate aged wae gillnet surveys
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("./data/Fish_Climate_Filtering.csv") %>% 
  filter(metric == "adult_abundance" & state == "Minnesota" & 
           species_1 == "walleye") %>%
  mutate(month_min = 4) %>%
  filter(sampling_method != "Standard 3/4-in mesh, double frame trap net sets") %>%
  rename(survey.type = survey_type,
         sampling.method = sampling_method,
         area.group = area_group) %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -target_species_1,
         -target_species_2,
         -survey_type_effort)

#selects fish level filters to be applied in the second filtering step
filter_table_fish <- 
  read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Minnesota" & 
           species_1 == "walleye") %>% 
  #fish level parameters of interest 
  select(state) %>% 
  distinct() 

good_surveys <- aged %>% 
  right_join(filter_table_effort, by = c("state",
                                        "survey.type",
                                        "sampling.method",
                                        "area.group")) %>% 
  distinct(lake.id,
           date,
           survey.type,
           sampling.method,
           total.effort.1,
           effort.units,
           area.group,
           effort_min,
           month_min,
           month_max, 
           alk) %>% 
  filter(total.effort.1 >= effort_min,
         month(date) >= month_min,
         month(date) <= month_max,
         alk == "year" # Only include fish that were aged at a lake
         ) %>% 
  collect()

adult_walleye_age <- aged %>% 
  right_join(good_surveys) %>% 
  #filter for fish level items (length, etc)
  right_join(filter_table_fish) %>% 
  collect() %>%
  mutate(survey_id = str_c(lake.id, year, total.effort.1)) %>%
  select(lake.name, lake.id, nhdhr.id, date, year,
         total.effort.1, effort.ident, survey.type, 
         sampling.method, total.effort.1, survey_id, species, est.age)
  

long <- adult_walleye_age %>% distinct(lake.id, year) %>% group_by(lake.id) %>% 
  summarize(n = n()) #%>% filter(n > 4)


id = unique(adult_walleye_age$lake.id)
effort_id = unique(adult_walleye_age$survey_id)
```


# filter for appropriate MISSING FISH from surveys
Some fish are not lengthed during a survey, these fish are not aged and thus not in the HALK data. I can get these fish from the MW hive, randomly sample from the distribution of ages (of that survey) and then add them back into the dataset
```{r}
mw_data <- open_dataset("./data/hive", partitioning = c("state"))

filter_table_effort_hive <- filter_table_effort %>% 
  rename(survey_type = survey.type, 
         sampling_method = sampling.method, 
         area_group = area.group)

good_surveys_hive <- mw_data %>% 
  filter(state == "Minnesota") %>%
  right_join(filter_table_effort_hive, by = c("state",
                                        "survey_type",
                                        "sampling_method",
                                        "area_group")) %>% 
  distinct(lake_id,
           survey_type,
           sampling_method,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units,
           year,
           area_group,
           effort_min,
           month,
           month_min,
           month_max) %>% 
  collect() %>% 
  filter(total_effort_1 >= effort_min,
         month >= month_min,
         month <= month_max,
         lake_id %in% id) %>%
  mutate(survey_id = str_c(lake_id, year, total_effort_1))



adult_walleye <- mw_data %>% 
  filter(state == "Minnesota") %>%
  mutate(survey_id = str_c(lake_id, year, total_effort_1)) %>%
  filter(survey_id %in% effort_id) %>%
  filter(species_1 == "walleye") %>%
  right_join(good_surveys_hive) %>% 
  collect() %>%
  filter(is.na(length_1)) %>%
  dplyr::select(lake_name, lake_id, nhdhr_id, date_survey, year, 
                total_effort_1, total_effort_ident, survey_type, 
                sampling_method, survey_id, species_1, age_class) %>%
  mutate(est.age = ifelse(age_class == "young_of_year", 0, NA)) %>%
  select(!age_class)
  
unaged_id <- unique(adult_walleye$survey_id)

# For each walleye that is unaged
for ( i in 1:nrow(adult_walleye)){
  # Get the ages of the walleye from the survey of the unaged walleye
  walleye_visit = adult_walleye_age %>% 
    filter(survey_id == adult_walleye$survey_id[i]) %>%
    select(est.age)
  # If the fish is unaged (isn't age-0) and the survey exists in the halk data
  if (is.na(adult_walleye$est.age[i]) && nrow(walleye_visit) > 0){
    adult_walleye$est.age[i]  <- sample(walleye_visit$est.age, size = 1)
  }
}

colnames(adult_walleye) <- colnames(adult_walleye_age)
adult_walleye_age <- rbind(adult_walleye, adult_walleye_age)

#creating cpue for species of interest
adult_walleye_cpue_aged <- adult_walleye_age %>% 
  #grouping by variables to get survey fish count 
  group_by(lake.id,
           date,
           survey.type,
           sampling.method,
           total.effort.1, est.age) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species == 'walleye')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total.effort.1)) %>%
  dplyr::filter(species == 'walleye') %>%
  ungroup() %>%
  distinct(lake.name, lake.id, date, year, total.effort.1,  
           survey.type, sampling.method, total.effort.1,
           est.age, count, cpue, .keep_all = T) %>%
  collect() %>%
  select(lake.name, lake.id, date, year, total.effort.1, 
         survey.type, sampling.method, total.effort.1, effort.ident, 
         est.age, count, cpue) %>%
  mutate(birth.year = year - est.age, 
         julian_day = yday(date))
```


# Add in ages that are not captured when a survey occured
 If we sample a lake and catch only 9 year old fish that means age 1, 2, 3 etc year old fish should be zeroes 
```{r}
walleye_anomaly <- adult_walleye_cpue_aged %>%
  mutate(lake.id = mnsentinellakes::fixlakeid(lake.id),
         lake.id = as.factor(lake.id),
         year_f = as.factor(year),
         jd = lubridate::yday(date),
         survey_id = str_c(lake.id, date)) %>%
  filter(est.age <= 25) %>%
  filter(lake.id != 69129100 | survey.type != "HISTORICAL")  

surveys <- unique(walleye_anomaly$survey_id)

missing.ages <- walleye_anomaly[0,]

for (i in 1:length(surveys)) {
  temp <- walleye_anomaly %>% dplyr::filter(survey_id == surveys[i])
  early = min(temp$birth.year); late = max(temp$birth.year)
  diff = late - early
  missing = 1:max(temp$est.age) %in% temp$est.age
  for (j in 1:length(missing)){
    if(missing[j] == F){
      missing.year = temp[1,]
      missing.year$est.age = j
      missing.year$count = 0; missing.year$cpue = 0
      missing.ages <- rbind(missing.ages, missing.year)
    }#if
  }# for j
}# for i

missing.ages %>% 
  group_by(survey_id, est.age) %>% count() %>% filter(n>1)

# Only select 3-year old fish
walleye_anomaly_ages <-  rbind(walleye_anomaly, missing.ages) %>%
  mutate(lake.id = as.factor(mnsentinellakes::fixlakeid(lake.id))) %>%
  filter(est.age == 3)

walleye_anomaly_ages %>% 
  group_by(survey_id, est.age) %>% count() %>% filter(n>1)


```


# Add covariate data
```{r}
st_crs(mn_lakes) <- 26915
mn_gn_lakes <- mn_lakes %>% dplyr::filter(dowlknum %in% walleye_anomaly_ages$lake.id)

mn_gn_lakes <- st_as_sf(x = mn_lakes,                         
                  coords = c("x", "y"),
                  crs = "+proj=utm +zone=15")
mn_gn_lakes = st_transform(mn_lakes, crs = "+proj=longlat +datum=WGS84")
mn_gn_lakes$x <- st_coordinates(st_centroid(st_as_sf(mn_gn_lakes)))[,"X"]
mn_gn_lakes$y <- st_coordinates(st_centroid(st_as_sf(mn_gn_lakes)))[,"Y"]

mn_gn_lakes_clean <- as.data.frame(mn_gn_lakes) %>% 
  select("dowlknum", "pw_basin_n","acres","wb_class", "shore_mi", "x", "y") %>%
  dplyr::filter(pw_basin_n != "Mississippi River - Pepin (WI)") %>%
  rename(lake.id = dowlknum) %>% distinct(lake.id, .keep_all = T)

gn_spatial <- merge(x = walleye_anomaly_ages, y = mn_gn_lakes_clean, 
                    by = "lake.id", all.x = T)


stocking_baby <- read.csv("./data/mn_wae_stocking.csv") %>% 
  rename(lake_id = DOW, year = Year) %>%
  mutate(lake_id = ifelse(lake_id == "04003500", "04003501", lake_id),
         recruit.year = year + 3) %>%
  select(lake_id, recruit.year, FRY, FGL)

stocking.yrl <- read.csv("./data/mn_wae_stocking.csv") %>% 
  rename(lake.id = DOW, year = Year) %>%
  mutate(recruit.year = year + 2) %>%
  select(lake.id, recruit.year, YRL)

gn_stocking <- merge(x = gn_spatial, y = stocking_baby, 
                     by.x = c("lake.id", "year"), 
                     by.y = c("lake_id", "recruit.year"),
                     all.x = T) %>%
  merge(stocking.yrl, 
        by.x = c("lake.id", "year"), 
        by.y = c("lake.id", "recruit.year"),
        all.x = T) %>%
  mutate(across(FRY:YRL, ~ ifelse(is.na(.), 0, .))) 


ais <- readxl::read_excel("./data/infested-waters.xlsx") %>% 
  dplyr::filter(ais == "zebra mussel") %>% select(!Column1) %>%
  dplyr::filter(dow != "NA", dow != "na") %>%
  dplyr::filter(!grepl("none", dow) ) %>%
  mutate(lake_id = fixlakeid(dow),
         lake_id = str_pad(lake_id, width = 8, side = "right", pad = "0")) %>%
  dplyr::rename(year.listed = year) %>%
  select(lake_id, year.listed, year_confirmed, ais) %>% 
  filter(year.listed <= 2022)

gn_ais <- merge(x = gn_stocking, y = ais, 
                         by.x = "lake.id", by.y = "lake_id",
                all.x = T) %>%
  mutate(ais = ifelse(is.na(ais), 0, 1),
         year_confirmed = as.numeric(year.listed),
         year_confirmed = ifelse(is.na(year_confirmed), NA, year_confirmed),
         infested = ifelse(year_confirmed <= year, 1, 0),
         ) %>%
  dplyr::select(!c(ais))
```

# Add in GDD
```{r}
current_temps <- read_feather("./data/lake_temperature_metrics_GLM_NLDAS.feather") %>% 
  dplyr::select(site_id, year, gdd_wtr_0c, gdd_wtr_5c, gdd_wtr_10c)

average_temps <- current_temps %>% 
  group_by(site_id) %>% 
  summarize(median_gdd_0 = median(gdd_wtr_0c), median_gdd_5 = median(gdd_wtr_5c))

dow_year <- gn_ais %>% distinct(lake.id, year)

dow_nhdhr <- readRDS("./data/mndow_nhdhr_xwalk.rds") %>% rename(lake_id = MNDOW_ID) %>%
  mutate(lake_id = gsub("^.{0,6}", "", lake_id))

missing_ids <- data.frame(
  lake_id = c("69069400", "69060800", "38053200", "38021100", "16063300"), 
  site_id = c("nhdhr_120019354", "nhdhr_80993113", 
               "nhdhr_72d19d48-754d-459b-a1f2-b1cfd8993b06", "nhdhr_80993749", 
               "nhdhr_120019082"))

dow_nhdhr = dow_nhdhr %>% rows_update(y = missing_ids, by = c("lake_id"))

surveys_no_temp <- merge(gn_ais, dow_nhdhr, 
                         by.x = c("lake.id"),
                         by.y = c("lake_id"), all.x = T)

filtered_temp <- current_temps %>% 
  dplyr::filter(site_id %in% surveys_no_temp$site_id) %>%
  rename(gdd_wtr_0c_year3 = gdd_wtr_0c,
         gdd_wtr_5c_year3 = gdd_wtr_5c,
         gdd_wtr_10c_year3 = gdd_wtr_10c)

temp_year2 <- current_temps %>% 
  dplyr::filter(site_id %in% surveys_no_temp$site_id) %>%
  mutate(year = year+1) %>%
  rename(gdd_wtr_0c_year2 = gdd_wtr_0c,
         gdd_wtr_5c_year2 = gdd_wtr_5c,
         gdd_wtr_10c_year2 = gdd_wtr_10c)

temp_year1 <- current_temps %>% 
  dplyr::filter(site_id %in% surveys_no_temp$site_id) %>%
  mutate(year = year+2) %>%
  rename(gdd_wtr_0c_year1 = gdd_wtr_0c,
         gdd_wtr_5c_year1 = gdd_wtr_5c,
         gdd_wtr_10c_year1 = gdd_wtr_10c)

temp_year0 <- current_temps %>% 
  dplyr::filter(site_id %in% surveys_no_temp$site_id) %>%
  mutate(year = year + 3) %>%
  rename(gdd_wtr_0c_year0 = gdd_wtr_0c,
         gdd_wtr_5c_year0 = gdd_wtr_5c,
         gdd_wtr_10c_year0 = gdd_wtr_10c)

surveys_ef <- merge(x = surveys_no_temp , y = filtered_temp, 
                    by = c("site_id", "year"), all.x = T) %>%
  merge(y = temp_year2, by = c("site_id", "year"), all.x = T) %>%
  merge(y = temp_year1, by = c("site_id", "year"), all.x = T) %>%
  merge(y = temp_year0, by = c("site_id", "year"), all.x = T) %>%
  merge(y = average_temps, by = c("site_id"), all.x = T)
```



```{r}
year_3_fish <- surveys_ef %>% filter(year < 2022) %>%
  mutate(year.listed = ifelse(is.na(year.listed), 0, year.listed),
         year_confirmed = ifelse(is.na(year_confirmed), 0, year_confirmed),
         infested = ifelse(is.na(infested), 0, infested)) %>%
  drop_na() %>%
  select(!site_id)
  

#write.csv(year_3_fish, "./cleaned_data/year_3_fish.csv", row.names = F)
```


# remove big objects
```{r}
rm(adult_walleye); rm(adult_walleye_age); rm(good_surveys); rm(good_surveys_hive); 
rm(mn_lakes): rm(adult_walleye_cpue_aged)
```
