---
title: "Figures"
output: html_document
date: "2025-12-12"
---

```{r setup, include=FALSE}
#remotes::install_github("https://github.com/mikemeredith/wiqid")
library(tidyverse)
library(ggthemes)
library(viridis)
library(ggspatial)
library(cowplot)
library(grid)
library(data.table)
library(bayesplot)

EF_Final <- read_csv("./cleaned_data/ef_for_model.csv")
GN_Final <- read_csv("./cleaned_data/gn_for_model.csv") 


stocking <- read.csv("./data/mn_wae_stocking.csv") %>% 
  rename(lake_id = DOW, year = Year) %>%
  mutate(lake_id = ifelse(lake_id == "04003500", "04003501", lake_id)) %>%
  select(lake_id, year, FRY, FGL, fry.pa, fgl.pa) %>% 
  filter(lake_id %in% EF_Final$lake_id)
```

# Lake stats
```{r}
length(unique(EF_Final$lake_id))
ef_number <- EF_Final %>% group_by(lake_id) %>% count()
median(ef_number$n)
# Number invaded lakes
EF_Final %>% distinct(lake_id, .keep_all = T) %>% summarize(sum = sum(treatment))



length(unique(GN_Final$lake.id))
gn_number <- GN_Final %>% group_by(lake.id) %>% count()
median(gn_number$n)
# Number invaded lakes
GN_Final %>% distinct(lake.id, .keep_all = T) %>% summarize(sum = sum(treatment))
```


```{r}
out_gn <- readRDS(file="./models/zipnb_gn_model_brms.rds" )
out_ef <- readRDS(file='./models/nb_ef_model_brms.rds')
summary(out_gn)
summary(out_ef)

# Convert mcmc draws to matrix
posterior_gn_matrix <- as.matrix(out_gn)
posterior_gn <- as.data.table(posterior_gn_matrix)[ ,c(1:13),] #take only the covariate columns

# Convert mcmc draws to matrix
posterior_ef_matrix <- as.matrix(out_ef)
posterior_ef <- as.data.table(posterior_ef_matrix)[ ,c(1:13),] #take only the first 13 columns
```


# Figure 1

Map showing locations of lakes and whether or not they have zebra mussels 
```{r Figure 1, warning = FALSE, message = FALSE}
MN <- map_data("state")%>% filter(region=="minnesota")

# walleye lakes, ZM in red, rest in blue
EF_Lakes_ZM <- 
  ggplot(data = MN, mapping = aes(x = long, y = lat, group=group))+
  coord_fixed(1.3) +
  geom_polygon(fill = "white", color = "black", size = 1.5)+
  geom_point(data = EF_Final, aes(x= x, y = y, group=25, color=as.factor(treatment),
                                  shape = as.factor(treatment)),alpha = 0.15, size = 4) +
  scale_color_manual(name = "ZM Invasion Status", labels = c("Uninvaded", "ZM Invaded"),
                     values = c("#00BFC4", "#F8766D")) +  
  scale_shape_manual(name = "ZM Invasion Status", labels = c("Uninvaded", "ZM Invaded"), values = c(16, 15)) +
  guides(color = guide_legend(override.aes = list(alpha = .5)), shape = guide_legend(override.aes = list(alpha = .5))) +
  labs(color = "ZM Invasion Status",x= "Longitude", y ="Latitude")+
  theme_void() +
  theme(legend.position = "bottom", legend.box.background = element_rect(color = "gray85"),
        legend.box.margin = margin(t = 2, r = 4, b = 2, l = 4),
        legend.key = element_rect(color = "gray90"),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 18, face = "bold"),
        legend.key.size = unit(1.5, "cm"))

EF_Lakes_ZM
```

```{r Figure 1B, warning = FALSE, message = FALSE}
MN <- map_data("state")%>% filter(region=="minnesota")

# walleye lakes, ZM in red, rest in blue
GN_Lakes_ZM <- ggplot(data = MN, mapping = aes(x = long, y = lat, group=group))+
  coord_fixed(1.3) +
  geom_polygon(color= "black", fill = "white", size = 1.5)+
  geom_point(data=GN_Final, aes(x= x, y = y, group=25,  color=as.factor(treatment),
                                      shape = as.factor(treatment)),alpha = 0.15, size = 4) +
  scale_color_manual(name = "ZM Invasion Status", labels = c("Uninvaded", "ZM Invaded"),
                     values = c("#00BFC4", "#F8766D")) +
  scale_shape_manual(name = "ZM Invasion Status", labels = c("Uninvaded", "ZM Invaded"), values = c(16, 15))+
  labs(color = "ZM Invasion Status") +
  theme_void() +
  theme(legend.position = "none")

GN_Lakes_ZM
```

Bounding box and North America Map for inset
```{r country}
canadamap <- map_data("world", c("Canada")) 

mexicomap <- map_data("world", c("Mexico"))

all_states <- map_data("state")

lakes <- map_data("lakes") #pulls in great lakes and large lakes

states <- subset(all_states, region %in% c("minnesota"))

xlim <- range(states$long) + c(-.3, 0.1) 
ylim <- range(states$lat) + c(-0.7, .65)
# the +/- 0.3 creates an offset so the bounding box is not counting off MN

N_AM_map <- ggplot() +  
  geom_map(data = canadamap, aes(map_id = region, group = subregion),  
           map = canadamap, color = "black", fill = "grey90") +
  # creates a map of Canada
  geom_map(data = mexicomap, aes(map_id = region, group = subregion),  
           map = mexicomap, color = "black", fill = "grey90")+
  #creates a map of Mexico
  geom_map(data = all_states, aes(map_id = region), 
           map = all_states, color = "black", fill = "grey90") +
  # Creates a map of US states with the existing Canada map
  geom_map(data = lakes,aes(map_id = region, group = subregion),
           map = lakes, color = "black", fill = "lightskyblue2")+
  #makes the lines around the great lakes cleaner, and can fill them with blue
  geom_rect(aes(xmin = xlim[1], xmax = xlim[2], ymin = ylim[1], ymax = ylim[2]), 
            colour = "red", fill = NA, lwd = 1) + 
  # creates a red rectangle around the y- and x-lim
  coord_map(xlim = c(-125, -66), ylim = c(25, 50)) +
  # defines the cutoff/edges of the map
  labs(x = NULL, y = NULL) + theme_bw() + theme(panel.grid = element_blank())+
  theme(panel.background = element_rect(fill = "lightskyblue2", color = "black"),
        axis.text.x = element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank()) #remove y axis ticks
```


```{r inset}
Final_Inset_Map <- ggdraw()+
  draw_plot(EF_Lakes_ZM +
              annotation_north_arrow(location = "br", 
                         which_north = "grid",
                         pad_x = unit(0.5, "in"),
                         pad_y = unit(0.5, "in"),
                         style = north_arrow_orienteering,
                         width = unit(.9, "cm"),
                         height = unit(.9, "cm")))+ #adds scale bar
  draw_plot(N_AM_map, x = 0.7, y = 0.3, width = .3, height = .25)

Final_Inset_Map

plot_grid(Final_Inset_Map, GN_Lakes_ZM, labels = c("A", "B"), label_size = 20)
ggsave("./figures/Figure_1.jpeg", dpi = 600, width = 18, height = 10)
```


# Figure 2 (BACI)

EF data
```{r}
NoZM_Pre <- as.numeric(exp(posterior_ef$b_Intercept))

#sd = sqrt(posterior[, "sigma_lake"]^2 + posterior[, "sigma_year"]^2)
#NoZM_Pre <- exp(posterior[,"b_0"] + rnorm(length(posterior[, "sigma_lake"]), 0, sd))

# Control pre
mean(NoZM_Pre)
quantile(NoZM_Pre, c(0.025, 0.975))

# Control Post
NoZM_Post <- as.numeric(exp(posterior_ef$b_Intercept + posterior_ef$b_post))
#NoZM_Post <- exp(posterior[,"b_0"] + posterior[,"b_post"] +  rnorm(length(posterior[, "sigma_lake"]), 0, sd))
mean(NoZM_Post)
quantile(NoZM_Post, c(0.025, 0.975))

# ZM Treatment Pre
ZM_Pre <- as.numeric(exp(posterior_ef$b_Intercept + posterior_ef$b_treatment))
#ZM_Pre <- exp(posterior[,"b_0"] + posterior[,"b_trt"] +  rnorm(length(posterior[, "sigma_lake"]), 0, sd))
mean(ZM_Pre)
quantile(ZM_Pre, c(0.025, 0.975))

# ZM Treatment Post
ZM_Post <- as.numeric(exp(posterior_ef$b_Intercept + posterior_ef$b_treatment + posterior_ef$b_post + posterior_ef$`b_post:treatment`))
#ZM_Post <- exp(posterior[,"b_0"] + posterior[,"b_trt"] +posterior[,"b_post_trt"] +  rnorm(length(posterior[, "sigma_lake"]), 0, sd))
mean(ZM_Post)
quantile(ZM_Post, c(0.025, 0.975))

## CALULATIONS

# Sentence 1 - results GN
ZM_BACI <- exp(posterior_ef$`b_post:treatment`)
mean(ZM_BACI)
quantile(ZM_BACI, c(0.025, 0.975))
(mean(ZM_BACI) - 1) * 100
(quantile(ZM_BACI, c(0.025, 0.975)) - 1) * 100


# Sentence 2 - Results GN
# Difference: ZM_Post - ZM_Pre
ZM_Diff_ef <- ZM_Post - ZM_Pre
print("Zebra mussel lakes")
mean(ZM_Diff_ef)
quantile(ZM_Diff_ef, c(0.025, 0.975))
# Percent change: ZM Post relative to ZM Pre
ZM_PctChange_ef <- (ZM_Post - ZM_Pre) / ZM_Pre * 100
print("Percent change in zebra mussel lakes")
mean(ZM_PctChange)
quantile(ZM_PctChange, c(0.025, 0.975))

# Sentence 3 - Results GN
# Difference: NO ZM_Post - NO ZM_Pre
NOZM_Diff_ef <- NoZM_Post - NoZM_Pre
print("Non zebra mussel lakes")
mean(NOZM_Diff_ef)
quantile(NOZM_Diff_ef, c(0.025, 0.975))
# Percent change: NO ZM Post relative to NO ZM Pre
NOZM_PctChange_ef <- (NoZM_Post - NoZM_Pre) / NoZM_Pre * 100
print("Percent change in non zebra mussel lakes")
mean(NOZM_PctChange_ef)
quantile(NOZM_PctChange_ef, c(0.025, 0.975))

# Sentence 4 - Results GN
# Difference between ZM and non ZM lakes
pre_diff <-  ZM_Pre - NoZM_Pre
print("Pre invasion lakes")
mean(pre_diff)
quantile(pre_diff, c(0.025, 0.975))
# Percent change: non-zm to zm lakes
pre_PctChange <- (ZM_Pre - NoZM_Pre) / NoZM_Pre * 100
mean(pre_PctChange)
quantile(pre_PctChange, c(0.025, 0.975))


BACI_Estimates1 <- tibble(
  Category = c("No_ZM_Pre", "No_ZM_Post", "ZM_Pre", "ZM_Post"),
  Mean = c(mean(NoZM_Pre), mean(NoZM_Post), mean(ZM_Pre), mean(ZM_Post)),
  Quant_2.5 = c(quantile(NoZM_Pre, 0.025), quantile(NoZM_Post, 0.025), quantile(ZM_Pre, 0.025), quantile(ZM_Post, 0.025)),
  Quant_97.5 = c(quantile(NoZM_Pre, 0.975), quantile(NoZM_Post, 0.975), quantile(ZM_Pre, 0.975), quantile(ZM_Post, 0.975)),
  ZM = c("Uninvaded", "Uninvaded", "ZM", "ZM"),
  Quant_83 = c(quantile(NoZM_Pre, 0.83), quantile(NoZM_Post, 0.83), quantile(ZM_Pre, 0.83), quantile(ZM_Post, 0.83)),
  Quant_17 = c(quantile(NoZM_Pre, 0.17), quantile(NoZM_Post, 0.17), quantile(ZM_Pre, 0.17), quantile(ZM_Post, 0.17)),
  pre_post = c("Pre", "Post", "Pre", "Post") %>% fct_relevel("Pre", "Post"))


BACI_PLOT_DAT_ef <- BACI_Estimates1 %>%
  mutate(Levels = fct_relevel(Category, "No_ZM_Pre", "No_ZM_Post", "ZM_Pre", "ZM_Post"))
```

GN data
```{r}
NoZM_Pre <- as.numeric(exp(posterior_gn$b_Intercept))

mean(NoZM_Pre)
quantile(NoZM_Pre, c(0.025, 0.975))

# Control Post
NoZM_Post <- as.numeric(exp(posterior_gn$b_Intercept + posterior_gn$b_post))
#NoZM_Post <- exp(posterior[,"b_0"] + posterior[,"b_post"] +  rnorm(length(posterior[, "sigma_lake"]), 0, sd))
mean(NoZM_Post)
quantile(NoZM_Post, c(0.025, 0.975))

# ZM Treatment Pre
ZM_Pre <- as.numeric(exp(posterior_gn$b_Intercept + posterior_gn$b_treatment))
#ZM_Pre <- exp(posterior[,"b_0"] + posterior[,"b_trt"] +  rnorm(length(posterior[, "sigma_lake"]), 0, sd))
mean(ZM_Pre)
quantile(ZM_Pre, c(0.025, 0.975))

# ZM Treatment Post
ZM_Post <- as.numeric(exp(posterior_gn$b_Intercept + posterior_gn$b_treatment + posterior_gn$b_post + posterior_gn$`b_post:treatment`))
#ZM_Post <- exp(posterior[,"b_0"] + posterior[,"b_trt"] +posterior[,"b_post_trt"] +  rnorm(length(posterior[, "sigma_lake"]), 0, sd))
mean(ZM_Post)
quantile(ZM_Post, c(0.025, 0.975))

## CALULATIONS
# Sentence 1 - results GN
ZM_BACI <- exp(posterior_gn$`b_post:treatment`)
mean(ZM_BACI)
quantile(ZM_BACI, c(0.025, 0.975))
(mean(ZM_BACI) - 1) * 100
(quantile(ZM_BACI, c(0.025, 0.975)) - 1) * 100

# Sentence 2 - Results GN
# Difference: ZM_Post - ZM_Pre
ZM_Diff_gn <- ZM_Post - ZM_Pre
print("Zebra mussel lakes")
mean(ZM_Diff_gn)
quantile(ZM_Diff_gn, c(0.025, 0.975))
# Percent change: ZM Post relative to ZM Pre
ZM_PctChange_gn <- (ZM_Post - ZM_Pre) / ZM_Pre * 100
print("Percent change in zebra mussel lakes")
mean(ZM_PctChange_gn)
quantile(ZM_PctChange_gn, c(0.025, 0.975))

# Sentence 3 - Results GN
# Difference: NO ZM_Post - NO ZM_Pre
NOZM_Diff_gn <- NoZM_Post - NoZM_Pre
print("Non zebra mussel lakes")
mean(NOZM_Diff_gn)
quantile(NOZM_Diff_gn, c(0.025, 0.975))
# Percent change: NO ZM Post relative to NO ZM Pre
NOZM_PctChange_gn <- (NoZM_Post - NoZM_Pre) / NoZM_Pre * 100
print("Percent change in non zebra mussel lakes")
mean(NOZM_PctChange_gn)
quantile(NOZM_PctChange_gn, c(0.025, 0.975))

# Sentence 4 - Results GN
# Difference between ZM and non ZM lakes
pre_diff <-  ZM_Pre - NoZM_Pre
print("Pre invasion lakes")
mean(pre_diff)
quantile(pre_diff, c(0.025, 0.975))
# Percent change: non-zm to zm lakes
pre_PctChange <- (ZM_Pre - NoZM_Pre) / NoZM_Pre * 100
mean(pre_PctChange)
quantile(pre_PctChange, c(0.025, 0.975))

BACI_Estimates1 <- tibble(
  Category = c("No_ZM_Pre", "No_ZM_Post", "ZM_Pre", "ZM_Post"),
  Mean = c(mean(NoZM_Pre), mean(NoZM_Post), mean(ZM_Pre), mean(ZM_Post)),
  Quant_2.5 = c(quantile(NoZM_Pre, 0.025), quantile(NoZM_Post, 0.025), quantile(ZM_Pre, 0.025), quantile(ZM_Post, 0.025)),
  Quant_97.5 = c(quantile(NoZM_Pre, 0.975), quantile(NoZM_Post, 0.975), quantile(ZM_Pre, 0.975), quantile(ZM_Post, 0.975)),
  ZM = c("Uninvaded", "Uninvaded", "ZM", "ZM"),
  Quant_83 = c(quantile(NoZM_Pre, 0.83), quantile(NoZM_Post, 0.83), quantile(ZM_Pre, 0.83), quantile(ZM_Post, 0.83)),
  Quant_17 = c(quantile(NoZM_Pre, 0.17), quantile(NoZM_Post, 0.17), quantile(ZM_Pre, 0.17), quantile(ZM_Post, 0.17)),
  pre_post = c("Pre", "Post", "Pre", "Post") %>% fct_relevel("Pre", "Post"))

BACI_PLOT_DAT_gn <- BACI_Estimates1 %>%
  mutate(Levels = fct_relevel(Category, "No_ZM_Pre", "No_ZM_Post", "ZM_Pre", "ZM_Post"))
```

Plot figure
```{r, warning = FALSE}
# Figure A
BACI_Figure_ef <- BACI_PLOT_DAT_ef %>% 
  ggplot(aes(x = pre_post, y = Mean, ymin = Quant_2.5, ymax = Quant_97.5)) +
  geom_pointrange(aes(color = ZM, shape = ZM), size = 1.5, position = position_dodge(0.25))+
  geom_linerange(aes(x = pre_post, ymin = Quant_17, ymax = Quant_83, color = ZM), 
                 position = position_dodge(0.25), size = 3) +
  geom_line(aes(x = pre_post, y = Mean, color = ZM, group = ZM), 
            position = position_dodge(0.25), linetype = "dotted", size = 1.5) + 
  labs(y = "Age-0 Walleye CPUE")+
  theme_bw() +
  scale_color_manual(name = "ZM Invasion Status", labels = c("Uninvaded", "ZM"),
                     values = c("#00BFC4", "#F8766D")) + 
  scale_shape_manual(name = "ZM Invasion Status", labels = c("Uninvaded", "ZM"), values = c(16, 15))+
  theme(axis.text.x = element_text(color = "black", size = 14),
        axis.title.y = element_text(color = "black", size = 14),
        axis.title.x = element_blank()) +
  theme(legend.position = c(.8, .875), legend.box.background = element_rect(color = "gray85"),
        legend.box.margin = margin(t = 2, r = 4, b = 2, l = 4),
        legend.key = element_rect(color = "gray90"),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12, face = "bold"),
        legend.key.size = unit(.6, "cm"))
BACI_Figure_ef

# Figure B
# Summarize posteriors
plot_dat <- tibble(category = c("Uninvaded", "ZM Invaded"),
                   Mean = c(mean(NOZM_PctChange_ef), mean(ZM_PctChange_ef)),
                   Q2.5 = c(quantile(NOZM_PctChange_ef, 0.025), quantile(ZM_PctChange_ef, 0.025)),
                   Q97.5 = c(quantile(NOZM_PctChange_ef, 0.975), quantile(ZM_PctChange_ef, 0.975)),
                   Q17 = c(quantile(NOZM_PctChange_ef, 0.17),  quantile(ZM_PctChange_ef, 0.17)),
                   Q83 = c(quantile(NOZM_PctChange_ef, 0.83),  quantile(ZM_PctChange_ef, 0.83)))

# Plot
ef_percent_change <- plot_dat %>% 
  mutate(category = fct_relevel(category, "ZM Invaded", "Uninvaded")) %>%
  ggplot(aes(x = Mean, y = category, color = category, shape = category)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_linerange(aes(xmin = Q2.5, xmax = Q97.5), linewidth = 1) +    
  geom_linerange(aes(xmin = Q17, xmax = Q83), linewidth = 3) +  
  geom_point(size = 4.5) + 
  scale_color_manual(name = "ZM Invasion Status", labels = c("ZM Invaded", "Uninvaded"),
                   values = c("#F8766D", "#00BFC4")) + 
  scale_shape_manual(name = "ZM Invasion Status", labels = c("ZM Invaded", "Uninvaded"), 
                   values = c(15, 16)) + 
  labs(x = "% Change From Pre - Post Time Periods (Age-0)", y = NULL) +
  theme_bw() +
  theme(axis.text = element_text(size = 14, color = "black"),
        axis.title.x = element_text(size = 14, color = "black"),
        legend.position = "none",
        plot.margin  = margin(r = 30, b = 5, l = 5))

# Figure C
BACI_Figure_gn <- BACI_PLOT_DAT_gn %>% 
  ggplot(aes(x = pre_post, y = Mean, ymin = Quant_2.5, ymax = Quant_97.5)) +
  geom_pointrange(aes(color = ZM, shape = ZM), size = 1.5, position = position_dodge(0.25))+
  geom_linerange(aes(x = pre_post, ymin = Quant_17, ymax = Quant_83, color = ZM), 
                 position = position_dodge(0.25), size = 3)+
  geom_line(aes(x = pre_post, y = Mean, color = ZM, group = ZM), 
            position = position_dodge(0.25), linetype = "dotted", size = 1.5) + 
  labs(y = "Age-3 Walleye CPUE")+
  theme_bw() +
  scale_color_manual(name = "ZM Invasion Status", labels = c("Uninvaded", "ZM"),
                     values = c("#00BFC4", "#F8766D")) + 
  scale_shape_manual(name = "ZM Invasion Status", labels = c("Uninvaded", "ZM"), values = c(16, 15))+
  theme(axis.text.x = element_text(color = "black", size = 14),
        axis.title.y = element_text(color = "black", size = 14),
        axis.title.x = element_blank(), 
        legend.position = "none")
BACI_Figure_gn

# Figure D
plot_dat_gn <- tibble(category = c("Uninvaded", "ZM Invaded"),
                   Mean = c(mean(NOZM_PctChange_gn), mean(ZM_PctChange_gn)),
                   Q2.5 = c(quantile(NOZM_PctChange_gn, 0.025), quantile(ZM_PctChange_gn, 0.025)),
                   Q97.5 = c(quantile(NOZM_PctChange_gn, 0.975), quantile(ZM_PctChange_gn, 0.975)),
                   Q17 = c(quantile(NOZM_PctChange_gn, 0.17),  quantile(ZM_PctChange_gn, 0.17)),
                   Q83 = c(quantile(NOZM_PctChange_gn, 0.83),  quantile(ZM_PctChange_gn, 0.83)))

# Plot
gn_percent_change <- plot_dat_gn %>% 
  mutate(category = fct_relevel(category, "ZM Invaded", "Uninvaded")) %>%
  ggplot(aes(x = Mean, y = category, color = category, shape = category)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_linerange(aes(xmin = Q2.5, xmax = Q97.5), linewidth = 1) +   
  geom_linerange(aes(xmin = Q17,  xmax = Q83), linewidth = 3) +    
  geom_point(size = 4.5) + 
  scale_color_manual(name = "ZM Invasion Status", labels = c("ZM Invaded", "Uninvaded"),
                   values = c("#F8766D", "#00BFC4")) + 
  scale_shape_manual(name = "ZM Invasion Status", labels = c("ZM Invaded", "Uninvaded"), 
                   values = c(15, 16)) + 
  labs(x = "% Change From Pre - Post Time Periods (Age-3)", y = NULL) +
  theme_bw() +
  theme(axis.text = element_text(size = 14, color = "black"),
        axis.title.x = element_text(size = 14, color = "black"),
        legend.position = "none",
        plot.margin  = margin(r = 30, b = 5, l = 5))

        
plot_grid(BACI_Figure_ef, ef_percent_change, BACI_Figure_gn, gn_percent_change,
          labels = c("A", "B", "C", "D"), ncol = 2)

ggsave("./figures/Figure_2.jpeg", height = 10, width = 10)
```


# Stocking effect (numbers)
```{r}
no_stocking_ef <- as.numeric(exp(posterior_ef$b_Intercept))
stocking_ef <- as.numeric(exp(posterior_ef$b_Intercept + posterior_ef$b_STOCKED))

# effects of covariates results numbers - EF
stock_ef <- exp(posterior_ef$`b_STOCKED`)
mean(stock_ef)
quantile(stock_ef, c(0.025, 0.975))
(mean(stock_ef) - 1) * 100
(quantile(stock_ef, c(0.025, 0.975)) - 1) * 100


# effects of covariates results numbers - GN
no_stocking_gn <- as.numeric(exp(posterior_gn$b_Intercept))
fry_stocking_gn <- as.numeric(exp(posterior_gn$b_Intercept + posterior_gn$b_fry))

fry_gn <- exp(posterior_gn$`b_fry`)
mean(fry_gn)
quantile(fry_gn, c(0.025, 0.975))
(mean(fry_gn) - 1) * 100
(quantile(fry_gn, c(0.025, 0.975)) - 1) * 100


fgl_stocking_gn <- as.numeric(exp(posterior_gn$b_Intercept + posterior_gn$b_fgl))

fgl_gn <- exp(posterior_gn$`b_fingerling`)
mean(fgl_gn)
quantile(fgl_gn, c(0.025, 0.975))
(mean(fgl_gn) - 1) * 100
(quantile(fgl_gn, c(0.025, 0.975)) - 1) * 100
```


# Figure 3 (Posterior Distributions)

```{r}
color_scheme_set(scheme = "blue")

post_dist_ef <- mcmc_areas(
    posterior_ef,
    pars = c("b_treatment", "b_post","b_STOCKED","b_post:treatment","b_lake_area_s","b_median_gdd_5_s","b_anomaly_gdd_5_s"),
    prob = 0.95, # 95% intervals
    prob_outer = 0.99, # 99%
    point_est = "median",
    area_method = "scaled height"
  ) + 
  geom_vline(xintercept = 0, color = "red", linetype = "dashed") + 
  scale_y_discrete(limit = c("b_treatment", "b_post","b_post:treatment","b_lake_area_s", "b_STOCKED",
                             "b_median_gdd_5_s","b_anomaly_gdd_5_s"),
                       labels = c("Zebra Mussels", "Post Invasion", "ZM * Post", "Lake Area",
                                  "Stocked", "GDD Anomaly\n(age 0)", "Lake median GDD"))+
  theme_classic() +
  theme(axis.text.x = element_text(color = "black"),
        axis.text.y = element_text(color = "black", size = 14) )

post_dist_gn <- mcmc_areas(
    posterior_gn,
    pars = c("b_treatment", "b_post", "b_post:treatment","b_lake_area_s", "b_fry1", "b_fingerling1", 
             "b_anomaly_gdd_5_s", "b_median_gdd_5_s"),
    prob = 0.95, # 95% intervals
    prob_outer = 0.99, # 99%
    point_est = "median",
    area_method = "scaled height"
  ) + 
  geom_vline(xintercept = 0, color = "red", linetype = "dashed") + 
  scale_y_discrete(limit = c("b_treatment", "b_post", "b_post:treatment","b_lake_area_s", "b_fry1", "b_fingerling1", 
                             "b_anomaly_gdd_5_s", "b_median_gdd_5_s"),
                   labels = c("Zebra Mussels", "Post Invasion", "ZM * Post", "Lake Area", 
                              "Stocked (Fry)", "Stocked (Fingerlings)", 
                              "GDD Anomaly\n(age 0 - 2)", "Lake Median GDD")) +
  theme_classic() +
  theme(axis.text.x = element_text(color = "black"),
        axis.text.y = element_text(color = "black", size = 14))

plot_grid(post_dist_ef, post_dist_gn, labels = c("A", "B"))


ggsave("./figures/Figure_3.jpeg", height = 10, width = 12, dpi = 600)
```


# Figure 4

Stocking Figure
```{r, warning = FALSE}
# get stocking estimates

not_stocked <- exp(posterior_ef$b_Intercept) 

stocked <- exp(posterior_ef$b_Intercept + posterior_ef$b_STOCKED)


stocking_estimates <- tibble(
  Stocking_Status = c("Not Stocked", "Stocked"),
  Mean = c(mean(not_stocked), mean(stocked)),
  Quant_2.5 = c(quantile(not_stocked, 0.025), quantile(stocked, 0.025)),
  Quant_97.5 = c(quantile(not_stocked, 0.975), quantile(stocked, 0.975)),
  Quant_83 = c(quantile(not_stocked, 0.83), quantile(stocked, 0.83)),
  Quant_17 = c(quantile(not_stocked, 0.17), quantile(stocked, 0.17)))


stocking_estimates_DAT <- stocking_estimates %>%
  mutate(Levels = fct_relevel(Stocking_Status, "Not Stocked", "Stocked"))



stocking_figure_ef <- ggplot(stocking_estimates_DAT , aes(x = Stocking_Status, y = Mean, ymin = Quant_2.5, ymax = Quant_97.5))+
  #geom_segment(aes(x = "Not Stocked", y = 4.189089, xend = "Stocked", yend = 11.138306), color = "grey85", linetype = "dashed")+
  geom_pointrange(aes(color = Stocking_Status),size = 1.5)+
  geom_linerange(aes(x = Levels, ymin = Quant_17, ymax = Quant_83, color = Stocking_Status), size = 4)+
  scale_discrete_manual(aesthetics = c("color", "fill"), values = c("#4DAF4A", "#984EA3"))+
  labs(y = "Age-0 Walleye CPUE")+
  theme_bw()+
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(color = "black", size = 14),
        axis.title.y = element_text(color = "black", size = 14),
        axis.title.x = element_blank())

# Get stocking estimates (gillnet)
not_stocked <- exp(posterior_gn$b_Intercept) 
stocked_fry <- exp(posterior_gn$b_Intercept + posterior_gn$b_fry1) 
stocked_fgl <- exp(posterior_gn$b_Intercept + posterior_gn$b_fingerling1)


stocking_estimates <- tibble(stocking_estimates <- tibble(stocking_estimates <- tibble(
  Stocking_Status = c("Not Stocked", "Fry Stocked", "Fingerling Stocked"),
  Mean = c(mean(not_stocked), mean(stocked_fry), mean(stocked_fgl)),
  Quant_2.5 = c(quantile(not_stocked, 0.025), quantile(stocked_fry, 0.025), quantile(stocked_fgl, 0.025)),
  Quant_97.5 = c(quantile(not_stocked, 0.975), quantile(stocked_fry, 0.975), quantile(stocked_fgl, 0.975)),
  Quant_83 = c(quantile(not_stocked, 0.83), quantile(stocked_fry, 0.83), quantile(stocked_fgl, 0.83)),
  Quant_17 = c(quantile(not_stocked, 0.17), quantile(stocked_fry, 0.17), quantile(stocked_fgl, 0.17)))))


stocking_estimates_gn <- stocking_estimates %>%
  mutate(Levels = fct_relevel(Stocking_Status, "Not Stocked", "Fry Stocked", "Fingerling Stocked"))


stocking_figure_gn <-ggplot(stocking_estimates_gn , aes(x = Stocking_Status, y = Mean, ymin = Quant_2.5, ymax = Quant_97.5))+
  geom_pointrange(aes(color = Stocking_Status), size = 1.5)+
  geom_linerange(aes(x = Levels, ymin = Quant_17, ymax = Quant_83, color = Stocking_Status), size = 4)+
  scale_discrete_manual(aesthetics = c("color", "fill"), values = c("#377EB8", "#984EA3", "#4DAF4A"))+
  labs(y = "Age-3 Walleye CPUE")+
  theme_bw() +
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(color = "black", size = 14),
        axis.title.y = element_text(color = "black", size = 14),
        axis.title.x = element_blank())
        

plot_grid(stocking_figure_ef, stocking_figure_gn, labels = c("A", "B"), rel_widths = c(.45, .55))
ggsave("./figures/Figure_4.jpeg", height = 8, width = 10)
```


# Figure 5

Environmental effects
GDD Anomaly (EF)
```{r}
# Begin predict code for first year GDD

b_gdd_sum <- posterior_ef$b_anomaly_gdd_5_s
b_0 <- posterior_ef$b_Intercept

sum <- EF_Final$anomaly_gdd_5 

lower_gdd = min(sum)
upper_gdd = max(sum)
gdd_range <- seq(lower_gdd, upper_gdd, by= 5) 

gdd_range_standardized <- wiqid::standardize2match(gdd_range, sum)

# Output storage DF
gdd.plot.anomaly <- data.frame(gdd = gdd_range, 
                        gdd_s = gdd_range_standardized, 
                        mean = numeric(length(gdd_range)), 
                        low = numeric(length(gdd_range)), high = numeric(length(gdd_range)))

for (i in 1:nrow(gdd.plot.anomaly)){
  pred.gdd = exp(b_0 + (b_gdd_sum * gdd.plot.anomaly[i, 2]))
  gdd.plot.anomaly[i, 3] = mean(pred.gdd)
  gdd.plot.anomaly[i, 4:5] = quantile(pred.gdd, c(.025, .975))
}

# Plot
gdd_plot_sum_ef <- gdd.plot.anomaly %>% 
  ggplot(aes(x = gdd)) +
  geom_line(aes(x= gdd, y = mean), size = 2)+
  geom_ribbon(aes(ymin = low, ymax = high), alpha= .2) +
  labs(x = "GDD Anomaly from Age-0 (base 5°C)", y = "Predicted Electrofishing CPUE")+
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) +
  theme_classic()

gdd_plot_sum_ef
```

GDD anomaly (GN)
```{r}
b_gdd_sum <- posterior_gn$b_anomaly_gdd_5_s
b_0 <- posterior_gn$b_Intercept

sum <- GN_Final$anomaly_gdd_5 

lower_gdd = min(sum)
upper_gdd = max(sum)
gdd_range <- seq(lower_gdd, upper_gdd, by= 5) 

gdd_range_standardized <- wiqid::standardize2match(gdd_range, sum)

# Output storage DF
gdd.plot.anomaly <- data.frame(gdd = gdd_range, 
                        gdd_s = gdd_range_standardized, 
                        mean = numeric(length(gdd_range)), 
                        low = numeric(length(gdd_range)), high = numeric(length(gdd_range)))

for (i in 1:nrow(gdd.plot.anomaly)){
  pred.gdd = exp(b_0 + (b_gdd_sum * gdd.plot.anomaly[i, 2]))
  gdd.plot.anomaly[i, 3] = mean(pred.gdd)
  gdd.plot.anomaly[i, 4:5] = quantile(pred.gdd, c(.025, .975))
}

# Plot
gdd_plot_sum_gn <- gdd.plot.anomaly %>% 
  ggplot(aes(x = gdd)) +
  geom_line(aes(x= gdd, y = mean), size = 2)+
  geom_ribbon(aes(ymin = low, ymax = high), alpha= .2) +
  labs(x = "GDD Anomaly from Age-0 (base 5°C)", y = "Predicted Gillnet CPUE")+
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) +
  theme_classic()


gdd_plot_sum_gn

```

Lake area (EF)
```{r}
b_area <- posterior_ef$b_lake_area
b_0 <- posterior_ef$b_Intercept

# Lake area range
lower_area = min(EF_Final$log_acres)
upper_area = max(EF_Final$log_acres)
log_lake_area_range <- seq(lower_area, upper_area, by= 0.1) #935 numbers

log_lake_area_range_standardized <- wiqid::standardize2match(log_lake_area_range, EF_Final$log_acres)

area.plot <- data.frame(area = log_lake_area_range, 
                        area_s = log_lake_area_range_standardized, 
                        mean = numeric(length(log_lake_area_range)), 
                        low = numeric(length(log_lake_area_range)), high = numeric(length(log_lake_area_range)))

for (i in 1:nrow(area.plot)){
  pred.area = exp(b_0 + (b_area *area.plot[i, 2]))
  area.plot[i, 3] = mean(pred.area)
  area.plot[i, 4:5] = quantile(pred.area, c(.025, .975))
}

area_plot_ef <- area.plot %>% 
  ggplot(aes(x = area))+
  geom_line(aes(x=area, y = mean), size = 2)+
  geom_ribbon(aes(ymin = low, ymax = high), alpha= .2)+
  labs(x = "log Lake Area (acres)", y = "Predicted Electrofishing CPUE")+
  theme_classic()

area_plot_ef
```

Lake area (GN)
```{r}
# Begin predict code

b_area <- posterior_gn$b_lake_area
b_0 <- posterior_gn$b_Intercept

# Lake area range
lower_area = min(GN_Final$log_acres)
upper_area = max(GN_Final$log_acres)
log_lake_area_range <- seq(lower_area, upper_area, by= 0.1) #935 numbers

log_lake_area_range_standardized <- wiqid::standardize2match(log_lake_area_range, GN_Final$log_acres)

area.plot <- data.frame(area = log_lake_area_range, 
                        area_s = log_lake_area_range_standardized, 
                        mean = numeric(length(log_lake_area_range)), 
                        low = numeric(length(log_lake_area_range)), high = numeric(length(log_lake_area_range)))

for (i in 1:nrow(area.plot)){
  pred.area = exp(b_0 + (b_area *area.plot[i, 2]))
  area.plot[i, 3] = mean(pred.area)
  area.plot[i, 4:5] = quantile(pred.area, c(.025, .975))
}

area_plot_gn <- area.plot %>% 
  ggplot(aes(x = area))+
  geom_line(aes(x=area, y = mean), size = 2)+
  geom_ribbon(aes(ymin = low, ymax = high), alpha= .2)+
  labs(x = "log Lake Area (acres)", y = "Predicted Gillnet CPUE")+
  theme_classic()

area_plot_gn
```

Median GDD, lake level (EF)
```{r}
# Begin predict code
b_gdd_LL <- posterior_ef$b_median_gdd_5_s
b_0 <- posterior_ef$b_Intercept

# Range of observed lake level GDD's (base 5)
lower_gdd = min(EF_Final$median_gdd_5)
upper_gdd = max(EF_Final$median_gdd_5)


gdd_LL_range <- seq(lower_gdd, upper_gdd, by= 5)
# Standardized the range
gdd_LL_range_standardized <- wiqid::standardize2match(gdd_LL_range, EF_Final$median_gdd_5)

# Output storage DF
gdd.plot <- data.frame(gdd = gdd_LL_range, 
                        gdd_s = gdd_LL_range_standardized, 
                        mean = numeric(length(gdd_LL_range)), 
                        low = numeric(length(gdd_LL_range)), high = numeric(length(gdd_LL_range)))

for (i in 1:nrow(gdd.plot)){
  pred.gdd = exp(b_0 + (b_gdd_LL * gdd.plot[i, 2]))
  gdd.plot[i, 3] = mean(pred.gdd)
  gdd.plot[i, 4:5] = quantile(pred.gdd, c(.025, .975))
}

# Plot
gdd_LL_plot_ef <- gdd.plot %>% 
  ggplot(aes(x = gdd))+
  geom_line(aes(x= gdd, y = mean), size = 2)+
  geom_ribbon(aes(ymin = low, ymax = high), alpha= .2)+
  labs(x = "Median Growing Degree Days", y = "Predicted Electrofishing CPUE")+
  theme_classic()
gdd_LL_plot_ef
```

Median GDD, lake level (GN)
```{r}
# Begin predict code
b_gdd_LL <- posterior_gn$b_median_gdd_5_s
b_0 <- posterior_gn$b_Intercept

# Range of observed lake level GDD's (base 5)
lower_gdd = min(GN_Final$median_gdd_5)
upper_gdd = max(GN_Final$median_gdd_5)


gdd_LL_range <- seq(lower_gdd, upper_gdd, by= 5)
# Standardized the range
gdd_LL_range_standardized <- wiqid::standardize2match(gdd_LL_range, GN_Final$median_gdd_5)

# Output storage DF
gdd.plot <- data.frame(gdd = gdd_LL_range, 
                        gdd_s = gdd_LL_range_standardized, 
                        mean = numeric(length(gdd_LL_range)), 
                        low = numeric(length(gdd_LL_range)), high = numeric(length(gdd_LL_range)))

for (i in 1:nrow(gdd.plot)){
  pred.gdd = exp(b_0 + (b_gdd_LL * gdd.plot[i, 2]))
  gdd.plot[i, 3] = mean(pred.gdd)
  gdd.plot[i, 4:5] = quantile(pred.gdd, c(.025, .975))
}

# Plot
gdd_LL_plot_gn <- gdd.plot %>% 
  ggplot(aes(x = gdd))+
  geom_line(aes(x= gdd, y = mean), size = 2)+
  geom_ribbon(aes(ymin = low, ymax = high), alpha= .2)+
  labs(x = "Median Growing Degree Days", y = "Predicted Gillnet CPUE")+
  theme_classic()
gdd_LL_plot_gn
```


# Combine all environmental figures
```{r}
Fig_5 <- plot_grid( gdd_LL_plot_ef, gdd_LL_plot_gn, gdd_plot_sum_ef, gdd_plot_sum_gn, area_plot_ef, area_plot_gn,
          labels = c("A", "B", "C", "D", "E", "F"), ncol = 2,
          label_size = 14)
Fig_5
ggsave("./figures/Figure_5.jpeg", height = 12, width = 10)
```

