---
title: "Figures"
author: "Holly Kundel"
date: "`r Sys.Date()`"
output: html_document
---


Load in required packages

```{r, message = FALSE, warning = FALSE}
library(rstan)
library(bayesplot)
library(dplyr)
library(lubridate)
library(tidyr)
library(stringr)
library(ggplot2)
library(forcats)
library(ggmcmc)
library(ggthemes)
library(readr)
library(readxl)
library(doBy)
library(RColorBrewer)
library(tibble)
library(tidybayes)
library(viridis)
library(ggspatial)
library(maps)
library(cowplot)
# library(ggsn)
library(grid)
library(data.table)
library(shinystan)
```

# Data
```{r, message=FALSE}
EF_Final <- read_csv("./cleaned_data/ef_for_model.csv")

out <- readRDS(file='./models/zinb_ef_model_brms.rds')

# launch_shinystan(out)


# Convert mcmc draws to matrix
posterior <- as.matrix(out)
glimpse(posterior)


posterior_df <- as.data.table(posterior)[ ,c(1:13),] #take only the first 13 columns

```




# Figure 1

Map showing locations of lakes and whether or not they have zebra mussels as of 2019

```{r Figure 1, warning = FALSE, message = FALSE}
MN <- map_data("state")%>% filter(region=="minnesota")
MN_Counties <- map_data("county")%>%
  filter(region == "minnesota")

# walleye lakes, ZM in red, rest in blue
EF_Lakes_ZM <- ggplot(data = MN_Counties, mapping = aes(x = long, y = lat, group=group))+
  coord_fixed(1.3) +
  geom_polygon(color= "gray85", fill = "gray90")+
  geom_point(data=EF_Final, aes(x= x, y = y, group=25, color=as.factor(treatment),
                                      shape = as.factor(treatment)),alpha = 0.3, size =3) +
  scale_color_manual(name = "ZM Invasion Status", labels = c("Uninvaded", "ZM Invaded"), values = c("#0571b0", "#ca0020"))+
  scale_shape_manual(name = "ZM Invasion Status", labels = c("Uninvaded", "ZM Invaded"), values = c(16, 15))+
  labs(color = "ZM Invasion Status",x= "Longitude", y ="Latitude")+
  theme_void()+
  theme(legend.position = "bottom",
        legend.box.background = element_rect(color = "gray85"),
        legend.box.margin = margin(t = 2, r = 4, b = 2, l = 4),
        legend.key = element_rect(color = "gray90"))+
  coord_sf(crs = 26991) #EPSG MN North https://epsg.io/26991

EF_Lakes_ZM
```

Bounding box and North America Map for inset
```{r country}
canadamap <- map_data("world", c("Canada")) 

mexicomap <- map_data("world", c("Mexico"))

all_states <- map_data("state")

lakes <- map_data("lakes") #pulls in great lakes and large lakes

states <- subset(all_states, region %in% c("minnesota"))

xlim <- range(states$long) + c(-.3, 0.1) 
ylim <- range(states$lat) + c(-0.7, .65)
# the +/- 0.3 creates an offset so the bounding box is not counting off MN

N_AM_map <- ggplot() +  
  geom_map(data = canadamap, aes(map_id = region, group = subregion),  
           map = canadamap, color = "black", fill = "grey90") +
  # creates a map of Canada
  geom_map(data = mexicomap, aes(map_id = region, group = subregion),  
           map = mexicomap, color = "black", fill = "grey90")+
  #creates a map of Mexico
  geom_map(data = all_states, aes(map_id = region), 
           map = all_states, color = "black", fill = "grey90") +
  # Creates a map of US states with the existing Canada map
  geom_map(data = lakes,aes(map_id = region, group = subregion),
           map = lakes, color = "black", fill = "lightskyblue2")+
  #makes the lines around the great lakes cleaner, and can fill them with blue
  geom_rect(aes(xmin = xlim[1], xmax = xlim[2], ymin = ylim[1], ymax = ylim[2]), 
            colour = "red", fill = NA, lwd = 1) + 
  # creates a red rectangle around the y- and x-lim
  coord_map(xlim = c(-125, -66), ylim = c(25, 50)) +
  # defines the cutoff/edges of the map
  labs(x = NULL, y = NULL) + theme_bw() + theme(panel.grid = element_blank())+
  theme(panel.background = element_rect(fill = "lightskyblue2", color = "black"),
        axis.text.x = element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank()) #remove y axis ticks
N_AM_map
```

Inset maps
```{r inset}
Final_Inset_Map <- ggdraw()+
  draw_plot(EF_Lakes_ZM +
              # ggsn::scalebar(data = MN_Counties, dist = 50, dist_unit = "km", st.size = 2, 
              #           transform = TRUE, model = 'WGS84', height = 0.01, location = "bottomright",
              #           box.color = c("black", "grey"))+
              annotation_north_arrow(location = "br", 
                         which_north = "grid",
                         pad_x = unit(0.7, "in"),
                         pad_y = unit(0.7, "in"),
                         style = north_arrow_orienteering,
                         width = unit(.7, "cm"),
                         height = unit(.7, "cm")))+ #adds scale bar
  draw_plot(N_AM_map, x = 0.7, y = 0.3, width = .2, height = .15)

Final_Inset_Map
ggsave("./figures/figure_1.jpeg", height = 10, width = 9)
```

# Figure 2

BACI Figure

Data
```{r}

NoZM_Pre <- as.numeric(exp(posterior_df$b_Intercept))


#sd = sqrt(posterior[, "sigma_lake"]^2 + posterior[, "sigma_year"]^2)
#NoZM_Pre <- exp(posterior[,"b_0"] + rnorm(length(posterior[, "sigma_lake"]), 0, sd))


mean(NoZM_Pre)
quantile(NoZM_Pre, c(0.025, 0.975))

# Control Post
NoZM_Post <- as.numeric(exp(posterior_df$b_Intercept + posterior_df$b_post))
#NoZM_Post <- exp(posterior[,"b_0"] + posterior[,"b_post"] +  rnorm(length(posterior[, "sigma_lake"]), 0, sd))
mean(NoZM_Post)
quantile(NoZM_Post, c(0.025, 0.975))

# ZM Treatment Pre
ZM_Pre <- as.numeric(exp(posterior_df$b_Intercept + posterior_df$b_treatment))
#ZM_Pre <- exp(posterior[,"b_0"] + posterior[,"b_trt"] +  rnorm(length(posterior[, "sigma_lake"]), 0, sd))
mean(ZM_Pre)
quantile(ZM_Pre, c(0.025, 0.975))

# ZM Treatment Post
ZM_Post <- as.numeric(exp(posterior_df$b_Intercept + posterior_df$b_treatment + posterior_df$b_post + posterior_df$`b_post:treatment`))
#ZM_Post <- exp(posterior[,"b_0"] + posterior[,"b_trt"] +posterior[,"b_post_trt"] +  rnorm(length(posterior[, "sigma_lake"]), 0, sd))
mean(ZM_Post)
quantile(ZM_Post, c(0.025, 0.975))

BACI_Estimates1 <- tibble(
  Category = c("No_ZM_Pre", "No_ZM_Post", "ZM_Pre", "ZM_Post"),
  Mean = c(mean(NoZM_Pre), mean(NoZM_Post), mean(ZM_Pre), mean(ZM_Post)),
  Quant_2.5 = c(quantile(NoZM_Pre, 0.025), quantile(NoZM_Post, 0.025), quantile(ZM_Pre, 0.025), quantile(ZM_Post, 0.025)),
  Quant_97.5 = c(quantile(NoZM_Pre, 0.975), quantile(NoZM_Post, 0.975), quantile(ZM_Pre, 0.975), quantile(ZM_Post, 0.975)),
  ZM = c("Uninvaded", "Uninvaded", "ZM", "ZM"),
  Quant_83 = c(quantile(NoZM_Pre, 0.83), quantile(NoZM_Post, 0.83), quantile(ZM_Pre, 0.83), quantile(ZM_Post, 0.83)),
  Quant_17 = c(quantile(NoZM_Pre, 0.17), quantile(NoZM_Post, 0.17), quantile(ZM_Pre, 0.17), quantile(ZM_Post, 0.17)))

BACI_PLOT_DAT <- BACI_Estimates1 %>%
  mutate(Levels = fct_relevel(Category, "No_ZM_Pre", "No_ZM_Post", "ZM_Pre", "ZM_Post"))
```


```{r, warning = FALSE}
brewer.pal(n = 2, "Set1") #E41A1C = red, #377EB8 = blue

BACI_Figure <-ggplot(BACI_PLOT_DAT , aes(x = Levels, y = Mean, ymin = Quant_2.5, ymax = Quant_97.5))+
  geom_pointrange(aes(color = ZM),size = 1.5)+
  geom_linerange(aes(x = Levels, ymin = Quant_17, ymax = Quant_83, color = ZM), size = 4)+
  #geom_segment(aes(x = "No_ZM_Pre", y = 4.189089, xend = "No_ZM_Post", yend = 4.464494), color = "#377EB8", linetype = "dashed")+
  #geom_segment(aes(x = "ZM_Pre", y = 9.186239, xend = "ZM_Post", yend = 5.449405), color ="#E41A1C", linetype = "dashed")+
  scale_color_brewer(palette = "Set1", direction = -1)+
  labs(y = "Mean Predicted Walleye Catch per Unit Effort")+
  theme_bw()+
  theme(legend.position = "none")+
  scale_x_discrete(limit = c("ZM_Pre", "ZM_Post", "No_ZM_Pre", "No_ZM_Post"),
                     labels = c("Pre ZM", "Post ZM", "Uninvaded Pre", "Uninvaded Post"))+
  theme(axis.text.x = element_text(family = "Arial", color = "black", size = 14),
        axis.title.x = element_blank())#,
        
BACI_Figure
ggsave("./zinb_figures/figure_2.jpeg", height = 10, width = 10)
```


# Figure 3
Posterior Distributions

```{r}
color_scheme_set(scheme = "blue")
post_dist_blue2 <- mcmc_areas(
    posterior,
    pars = c("b_treatment", "b_post","b_STOCKED","b_post:treatment","b_lake_area_s","b_median_gdd_5_s","b_anomaly_gdd_5_s"),
    prob = 0.95, # 95% intervals
    prob_outer = 0.99, # 99%
    point_est = "median",
    area_method = "scaled height"
  ) + 
  geom_vline(xintercept = 0, color = "red", linetype = "dashed") + 
  scale_y_discrete(limit = c("b_treatment", "b_post","b_post:treatment","b_lake_area_s", "b_STOCKED",
                             "b_median_gdd_5_s","b_anomaly_gdd_5_s"),
                       labels = c("Zebra Mussels", "Post Invasion", "ZM * Post", "Lake Area",
                                  "Stocked", "GDD (age 0)", "Lake median GDD"))+
  theme(axis.text.x = element_text(color = "black"),
        axis.text.y = element_text(color = "black", size = 14, family = "sans") )+
  theme_classic()

post_dist_blue2

ggsave("./zinb_figures/figure_3.jpeg", height = 10, width = 10)
```


# Figure 4

Stocking Figure
```{r, warning = FALSE}
# get stocking estimates

not_stocked <- exp(posterior_df$b_Intercept) 

stocked <- exp(posterior_df$b_Intercept + posterior_df$b_STOCKED)


stocking_estimates <- tibble(
  Stocking_Status = c("Not Stocked", "Stocked"),
  Mean = c(mean(not_stocked), mean(stocked)),
  Quant_2.5 = c(quantile(not_stocked, 0.025), quantile(stocked, 0.025)),
  Quant_97.5 = c(quantile(not_stocked, 0.975), quantile(stocked, 0.975)),
  Quant_83 = c(quantile(not_stocked, 0.83), quantile(stocked, 0.83)),
  Quant_17 = c(quantile(not_stocked, 0.17), quantile(stocked, 0.17)))


stocking_estimates_DAT <- stocking_estimates %>%
  mutate(Levels = fct_relevel(Stocking_Status, "Not Stocked", "Stocked"))


#get hex codes for colors
brewer.pal(n = 4, "Set1") #E41A1C = red, #377EB8 = blue, #4DAF4A = green, #984EA3 = purple

Stocking_Figure <-ggplot(stocking_estimates_DAT , aes(x = Stocking_Status, y = Mean, ymin = Quant_2.5, ymax = Quant_97.5))+
  #geom_segment(aes(x = "Not Stocked", y = 4.189089, xend = "Stocked", yend = 11.138306), color = "grey85", linetype = "dashed")+
  geom_pointrange(aes(color = Stocking_Status),size = 1.5)+
  geom_linerange(aes(x = Levels, ymin = Quant_17, ymax = Quant_83, color = Stocking_Status), size = 4)+
  scale_discrete_manual(aesthetics = c("color", "fill"), values = c("#4DAF4A", "#984EA3"))+
  labs(y = "Mean Predicted Walleye Catch per Unit Effort")+
  theme_bw()+
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(family = "Arial", color = "black", size = 14),
        axis.title.x = element_blank())#,
        
Stocking_Figure
ggsave("./zinb_figures/figure_4.jpeg", height = 10, width = 10)
```


# Figure 5

Environmental effects

```{r}
# Begin predict code for first three year GDD

b_gdd_sum <- posterior_df$b_anomaly_gdd_5_s
b_0 <- posterior_df$b_Intercept

sum <- EF_Final$anomaly_gdd_5 

lower_gdd = min(sum)
upper_gdd = max(sum)
gdd_range <- seq(lower_gdd, upper_gdd, by= 5) 

gdd_range_standardized <- wiqid::standardize2match(gdd_range, sum)

# Output storage DF
gdd.plot.anomaly <- data.frame(gdd = gdd_range, 
                        gdd_s = gdd_range_standardized, 
                        mean = numeric(length(gdd_range)), 
                        low = numeric(length(gdd_range)), high = numeric(length(gdd_range)))

for (i in 1:nrow(gdd.plot.anomaly)){
  pred.gdd = exp(b_0 + (b_gdd_sum * gdd.plot.anomaly[i, 2]))
  gdd.plot.anomaly[i, 3] = mean(pred.gdd)
  gdd.plot.anomaly[i, 4:5] = quantile(pred.gdd, c(.025, .975))
}

# Plot
gdd_Plot_sum <- gdd.plot.anomaly %>% 
  ggplot(aes(x = gdd)) +
  geom_line(aes(x= gdd, y = mean), size = 2)+
  geom_ribbon(aes(ymin = low, ymax = high), alpha= .2) +
  labs(x = "GDD Anomaly from Age-0 (base 5°C)", y = "Predicted CPUE")+
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) +
  theme_classic()


gdd_Plot_sum

```


Lake area
```{r}
# Begin predict code

b_area <- posterior_df$b_lake_area
b_0 <- posterior_df$b_Intercept

# Lake area range
lower_area = min(EF_Final$log_acres)
upper_area = max(EF_Final$log_acres)
log_lake_area_range <- seq(lower_area, upper_area, by= 0.1) #935 numbers

log_lake_area_range_standardized <- wiqid::standardize2match(log_lake_area_range, EF_Final$log_acres)

area.plot <- data.frame(area = log_lake_area_range, 
                        area_s = log_lake_area_range_standardized, 
                        mean = numeric(length(log_lake_area_range)), 
                        low = numeric(length(log_lake_area_range)), high = numeric(length(log_lake_area_range)))

for (i in 1:nrow(area.plot)){
  pred.area = exp(b_0 + (b_area *area.plot[i, 2]))
  area.plot[i, 3] = mean(pred.area)
  area.plot[i, 4:5] = quantile(pred.area, c(.025, .975))
}

area_Plot <- area.plot %>% 
  ggplot(aes(x = area))+
  geom_line(aes(x=area, y = mean), size = 2)+
  geom_ribbon(aes(ymin = low, ymax = high), alpha= .2)+
  labs(x = "log Lake Area (acres)", y = "Predicted CPUE")+
  theme_classic()

area_Plot
```


Median GDD (1993 - 2019), lake level
```{r, eval = F}
# Begin predict code
b_gdd_LL <- posterior_df$b_median_gdd_5_s
b_0 <- posterior_df$b_Intercept

# Range of observed lake level GDD's (base 5)
lower_gdd = min(EF_Final$median_gdd_5)
upper_gdd = max(EF_Final$median_gdd_5)


gdd_LL_range <- seq(lower_gdd, upper_gdd, by= 5)
# Standardized the range
gdd_LL_range_standardized <- wiqid::standardize2match(gdd_LL_range, EF_Final$median_gdd_5)

# Output storage DF
gdd.plot <- data.frame(gdd = gdd_LL_range, 
                        gdd_s = gdd_LL_range_standardized, 
                        mean = numeric(length(gdd_LL_range)), 
                        low = numeric(length(gdd_LL_range)), high = numeric(length(gdd_LL_range)))

for (i in 1:nrow(gdd.plot)){
  pred.gdd = exp(b_0 + (b_gdd_LL * gdd.plot[i, 2]))
  gdd.plot[i, 3] = mean(pred.gdd)
  gdd.plot[i, 4:5] = quantile(pred.gdd, c(.025, .975))
}

# Plot
gdd_LL_Plot <- gdd.plot %>% 
  ggplot(aes(x = gdd))+
  geom_line(aes(x= gdd, y = mean), size = 2)+
  geom_ribbon(aes(ymin = low, ymax = high), alpha= .2)+
  labs(x = "Median Growing Degree Days", y = "Predicted CPUE")+
  theme_classic()
gdd_LL_Plot
```



# Combine into one figure
```{r}
Fig_5 <- plot_grid(gdd_LL_Plot, gdd_Plot_sum,  area_Plot,
          labels = c("A", "B", "C"),
          label_size = 14)
Fig_5
ggsave("./zinb_figures/figure_5.jpeg", height = 10, width = 10)
```



# Summaries referenced in the manuscript, Table 1, and figures


Additional Data Summaries

Calculate percent change in pre vs post - zm
```{r}
#calculate percent change in pre-post
zm_diff = ZM_Pre - ZM_Post
zm_diff_prop = zm_diff/ZM_Pre

mean(zm_diff) #3.736835 walleye per hour
mean(zm_diff_prop) #40.67%
quantile(zm_diff_prop, c(0.025, 0.975))

uninvaded_diff = NoZM_Pre - NoZM_Post
mean(uninvaded_diff) #-0.2754049
uninvaded_prop = uninvaded_diff/NoZM_Pre
mean(uninvaded_prop)
quantile(uninvaded_prop, c(0.025, 0.975)) #-0.06585174 or 6.59% increase post
```

Effect of Stocking, stocking led to higher catch rates
```{r}
# Not stocked
Not_stocked <- exp(posterior[,"b_Intercept"]) 
mean(Not_stocked)
quantile(Not_stocked, c(0.025, 0.975)) 

#Stocked
stocked <- exp(posterior[,"b_Intercept"]+posterior[,"b_STOCKED"]) 
mean(stocked) 
quantile(stocked, c(0.025, 0.975)) 


#stocking effect
stocking_diff= stocked - Not_stocked
mean(stocking_diff) # 6.949217
quantile(stocking_diff, c(0.025, 0.975)) 

# percent change with stocking
stocking_prop = stocking_diff/Not_stocked
mean(stocking_prop) 
quantile(stocking_prop, c(0.025, 0.975)) 
```

#### JUNKYARD - Not up to date!
Basic Summary Info (Table 1)
```{r}
final.dat3 <- read_csv("R and stan code/Data/NN_EF_WAE_ZM_Model_MRVrerun.csv")

#surveys done in invaded and uninvaded lakes
Model_3_Summary <- final.dat3 %>%
  mutate(time = ifelse(!is.na(YEAR_INFESTED), YEAR_INFESTED, YEAR_INFESTED.y),
         post = ifelse(year > time, 1, 0))%>%
  mutate(Category = case_when(ZM_LAKE == 0 ~ "Uninvaded",
                                         ZM_LAKE == 1 & post == 0 ~ "PRE_ZM",
                                         ZM_LAKE == 1 & post == 1 ~ "POST_ZM"))%>%
  group_by(Category)%>%
  summarise(Total_Surveys = n()) #99 surveys post ZM, 223 surveys pre ZM, 1116 surveys uninvaded

#number of surveys done in each category
Model_3_Summary2 <- final.dat3 %>%
  mutate(time = ifelse(!is.na(YEAR_INFESTED), YEAR_INFESTED, YEAR_INFESTED.y),
         post = ifelse(year > time, 1, 0))%>%
  mutate(Category = case_when(ZM_LAKE == 0 & post == 0 ~ "Uninvaded_PRE",
                              ZM_LAKE == 0 & post == 1 ~ "Uninvaded_POST",
                                         ZM_LAKE == 1 & post == 0 ~ "PRE_ZM",
                                         ZM_LAKE == 1 & post == 1 ~ "POST_ZM"))%>%
  group_by(Category)%>%
  summarise(Total_Surveys = n()) # 99 post ZM, 223 pre zm, 181 uninvaded post, 935 uninvaded pre


# values of survey level covariates
Model_3_Summary_Survey_Level <- final.dat3 %>%
  mutate(time = ifelse(!is.na(YEAR_INFESTED), YEAR_INFESTED, YEAR_INFESTED.y),
         post = ifelse(year > time, 1, 0))%>%
  mutate(Category = case_when(ZM_LAKE == 0 & post == 0 ~ "Uninvaded_PRE",
                              ZM_LAKE == 0 & post == 1 ~ "Uninvaded_POST",
                                         ZM_LAKE == 1 & post == 0 ~ "PRE_ZM",
                                         ZM_LAKE == 1 & post == 1 ~ "POST_ZM"))%>%
  dplyr::select(SURVEY_YEAR, DD5, SURFACE_TEMP_F, AREA_HECTARES, MAX_DEPTH_M ,MEDIAN_93_19_GDD, TOTAL_CATCH,CPUE, YEAR_INFESTED , Category) %>%
  group_by(Category)%>%
  summarise_all(.funs = c(MEAN = mean.default, MAX = max, MIN = min), na.rm = TRUE) # %>%
  # select("Category", contains("CPUE")) #<- No worky
  

#write_csv(Model_3_Summary_Survey_Level, "Survey_Level_Summaries.csv")
# 
#values of lake level co-variates
Model_3_Summary_Lake_Level <- final.dat3 %>%
  mutate(time = ifelse(!is.na(YEAR_INFESTED), YEAR_INFESTED, YEAR_INFESTED.y),
         post = ifelse(year > time, 1, 0))%>%
  dplyr::select(36,37,38,44, 54,56, 61:63)%>%
  group_by(ZM_LAKE)%>%
  summarise_all(.funs = c(MEAN = mean, MAX = max, MIN = min), na.rm = TRUE)

# number of lakes in each category
Model_3_Summary_Lake <- final.dat3 %>%
  mutate(time = ifelse(!is.na(YEAR_INFESTED), YEAR_INFESTED, YEAR_INFESTED.y),
         post = ifelse(year > time, 1, 0))%>%
  mutate(Category = case_when(ZM_LAKE == 0 ~ "Uninvaded",
                                         ZM_LAKE == 1 & post == 0 ~ "PRE_ZM",
                                         ZM_LAKE == 1 & post == 1 ~ "POST_ZM"))%>%
  group_by(Category, DOW_KITTLE)%>%
  summarise(Total_Surveys = n())%>%
  group_by(Category)%>%
  summarise(Total_Lake_by_cat = n())
# 282 uninvaded lakes, 33 Pre-ZM, 33 Post-ZM lakes, 348 lakes total

# years surveys were done in each category
Model_3_Summary_Survey_Year <- final.dat3 %>%
  mutate(time = ifelse(!is.na(YEAR_INFESTED), YEAR_INFESTED, YEAR_INFESTED.y),
         post = ifelse(year > time, 1, 0))%>%
  mutate(Category = case_when(ZM_LAKE == 0 & post == 0 ~ "Uninvaded_PRE",
                              ZM_LAKE == 0 & post == 1 ~ "Uninvaded_POST",
                                         ZM_LAKE == 1 & post == 0 ~ "PRE_ZM",
                                         ZM_LAKE == 1 & post == 1 ~ "POST_ZM"))%>%
  dplyr::select(Category, SURVEY_YEAR)%>%
  group_by(Category)%>%
  summarise_all(.funs = c(MEAN = mean, MAX = max, MIN = min), na.rm = TRUE)
```

Posterior probabilities
```{r}
#GDD Lake Level average b_gdd

GDD_LL <- posterior_df%>%
  dplyr::select("b_gdd")%>%
  filter(b_gdd > 0)
nrow(GDD_LL)/75000 #0.7629467

#GDD survey Level average b_survey_gdd

GDD_LL <- posterior_df%>%
  dplyr::select(b_survey_gdd)%>%
  filter(b_survey_gdd > 0)
nrow(GDD_LL)/75000 # 0.14944

```
