---
title: "EF_WAE_ZM_Model1"
author: "Holly Kundel"
date: "11/15/2021"
output: html_document
---

This model is run using `ZM_LAKE` as zebra mussel invasion variable
- Only considered invaded if at least one EF survey was done the year of invasion or later
- No invasions after 2019 count (because no data post 2019)
- "0" indicates a lake is uninvaded (based on above criteria)
- "1" indicates a lake is invaded (based on above criteria)

Load in Required Packages
```{r}
#remove.packages(c("StanHeaders", "rstan"))
#install.packages("rstan", repos = "https://cloud.r-project.org/", dependencies = TRUE)
library("rstan") # observe startup messages
library(readr)
library(dplyr)
library(tidyr)
library(lubridate)
library(stringr)
library(arm)
library(doBy)
library(sf)
library(spData)
library(RANN)
library(ggplot2)
library(MASS)
library(bayesplot)
```

Load in Data
```{r, message=FALSE}
dat <- read_csv("./cleaned_data/yoy_catch_data.csv") %>% 
  mutate(survey_id = row_number(), 
         STOCKED = ifelse(FRY > 0,1,0))

nrow(dat[dat$catch==0,])/nrow(dat)
```
# Nearest Neighbor Code

- Assigning uninvaded lakes a "pseudo-invasion year" for BACI set up

Data setup
```{r, eval = F}
# Format dates
dat <- EF_WAE_ZM_data %>% 
  mutate(SAMPLING_START_DATE = mdy(SAMPLING_START_DATE),
         day = day(SAMPLING_START_DATE),
         month = month(SAMPLING_START_DATE),
         year = year(SAMPLING_START_DATE),
  )
```

Create Spatial sf Object
```{r}
# Create spatial sf object (4326 = WGS84)
dat.geo <- st_as_sf(dat, coords = c("x", "y"), crs = 4326)
```

Separate Invaded Lakes and Uninvaded Lakes

- Using column `ZM_LAKE` for this
- Only considered invaded if at least one EF survey was done the year of invasion or later
- No invasions after 2019 count (because no data post 2019)
- "0" indicates a lake is uninvaded (based on above criteria)
- "1" indicates a lake is invaded (based on above criteria)

```{r}
dat.geo.uninvaded <- dat.geo %>% 
  filter(year_confirmed == 0)
dim(dat.geo.uninvaded) #1116 surveys

dat.geo.invaded <- dat.geo %>% 
  filter(year_confirmed > 0)
dim(dat.geo.invaded) #322 surveys, adds up to correct total of 1438 surveys total
```

Get coordinate matrices
```{r}
uninvaded_coords <- do.call(rbind, st_geometry(dat.geo.uninvaded))
invaded_coords <- do.call(rbind, st_geometry(dat.geo.invaded))
invaded_coords <- cbind(invaded_coords, 1:nrow(invaded_coords)) # Index for identifying nearest lake

```

Find nearest neighbor
```{r}
closest <- nn2(invaded_coords[,1:2], uninvaded_coords,  k = 1, searchtype = "standard")
#str(closest)
#head(closest)
closest <- sapply(closest, cbind) 
```

Combine uninvaded data with nearest neighbor assignments
```{r}
dat.geo.uninvaded <- cbind(dat.geo.uninvaded, closest)
```

Give invaded lakes an identifier and select columns
```{r}
dat.geo.invaded$nn.idx <- 1:nrow(dat.geo.invaded)
dat.geo.invaded <- dat.geo.invaded %>% 
  dplyr::select(year_confirmed, nn.idx)
str(dat.geo.invaded)
head(dat.geo.invaded)

#str(dat.geo.uninvaded)
#str(dat.geo.invaded)
```

Merge uninvaded with invaded `YEAR_INFESTED` for use as "after" year in BACI (Before After Control Impact)
```{r}
dat.geo.invaded_merge <- st_set_geometry(dat.geo.invaded, NULL)

dat.geo.uninvaded.final <- dat.geo.uninvaded %>%
    left_join(dat.geo.invaded_merge, by = c("nn.idx"))
str(dat.geo.uninvaded.final)
summary(dat.geo.uninvaded.final)
# Bring back into full dataset
# Convert to dataframe
dat.geo.uninvaded.final.df <- st_set_geometry(dat.geo.uninvaded.final, NULL)
#class(dat.geo.uninvaded.final.df)
#str(dat.geo.uninvaded.final.df)

# Join dat.geo.uninvaded.final.df - want YEAR_INFESTED.y - with original data, dat
# Grab columns of interest
dat.geo.uninvaded.final.df <- dat.geo.uninvaded.final.df %>% 
  dplyr::select(survey_id, year_confirmed.y)
#dim(dat.geo.uninvaded.final.df)
#head(dat.geo.uninvaded.final.df)
#dim(dat)

# Final merge
final.dat3 <- dat %>%
    left_join(dat.geo.uninvaded.final.df, by = c("survey_id"))
#dim(final.dat3)
#head(final.dat3)
```

Export data with nearest neighbor
MAKE SURE EACH LAKE HAS A PRE AND A POST
```{r}
write.csv(final.dat3, "./cleaned_data/NN_EF_WAE_ZM_Model_CRrerun.csv", row.names = F)
```

Test Nearest Neighbor Data
```{r, eval =F}
# test
test1 <- final.dat3 %>%
   filter(ZM_LAKE == 0)
 head(test1)

# Export final data

# Check nearest neighbors
# Grab state boundaries from spData
us_states2 <- st_transform(us_states, crs = 4326)
# Rename column
colnames(us_states2)[2] <- "State"
selectStates <- c("Minnesota")
# Subset data for plotting
us_state_select <- us_states2[us_states2$State %in% selectStates, ]

# Subset for visualization
unique(dat.geo.uninvaded$nn.idx)
invaded.plot <- dat.geo.invaded %>% 
  filter(nn.idx == 307 | nn.idx == 309 |nn.idx == 45)

uninvaded.plot <- dat.geo.uninvaded %>% 
  filter(nn.idx == 307 | nn.idx == 309 |nn.idx == 45)

ggplot() + geom_sf(data = us_state_select, color = "gray30", lwd=1, fill="white") +
  geom_sf(data = invaded.plot, size=3.0, aes(color=as.factor(nn.idx))) +
  geom_point() +
  geom_sf(data = uninvaded.plot,size=0.3, aes(color=as.factor(nn.idx))) +
  geom_sf(data = dat.geo.invaded,size=0.3, color='black') +
  labs(title="", y="Latitude", x="Longitude", color = "Lakes") 
ggsave("Figs/NN_ZM_map.png", height=5, width=5, units="in")
```

# Prepare to run model

Load more libraries
```{r}
# if you need to install stan run: "install.packages("rstan", repos = "https://cloud.r-project.org/", dependencies = TRUE)"
#library(StanHeaders)
library(rstan)
library(bayesplot)
library(lme4)
library(ggmcmc) # Use ggmcmc package for assessing convergence
library(gridExtra)
library(ggthemes)
library(coda)
library(forcats)
```

Load in Data Again 
- could also use `final.dat` from above
```{r}
dat <- read.csv('./cleaned_data/NN_EF_WAE_ZM_Model_CRrerun.csv') %>%
  mutate(ZM_LAKE = ifelse(year_confirmed == 0, 0, 1))
#dim(dat)
#head(dat)
```

Create BACI design variables
```{r}
# "Treatment" is lake-level invasion status, ZM (1 = ZM; 0 otherwise)
# Before/After time variable ("post" = 1 if sampled after invasion; 0 if before)
# Below, "year" is year of sampling and time is year infested
dat <- dat %>% 
  mutate(time = ifelse(year_confirmed == 0, year_confirmed.y, year_confirmed),
         post = ifelse(year > time, 1, 0),
         treatment = ZM_LAKE # USING ZM_LAKE_HERE
         )%>%
  mutate(INVASION_STATUS_NEW = case_when(ZM_LAKE == 0 ~ "Uninvaded",
                                         ZM_LAKE == 1 & post == 0 ~ "PRE_ZM",
                                         ZM_LAKE == 1 & post == 1 ~ "POST_ZM"))

#head(dat)
#summary(dat)

#check data
dat %>%
  group_by(ZM_LAKE) %>%  
  count(post, ZM_LAKE) %>%
  ungroup() %>% 
  mutate(prop = prop.table(n))

###new addition of data filtering#######
#ensure at least one year pre and post
test = dat %>% 
  group_by(lake_id, post, ZM_LAKE) %>% 
  summarise(count=n()) %>% 
  filter(count > 0) %>% ungroup() %>% 
  group_by(lake_id, ZM_LAKE) %>% 
  summarise(pre.post.test = n()) %>% 
  filter(pre.post.test==2)


dat <- dat %>% filter(lake_id %in% test$lake_id)
```

Prep for model
```{r}
# Create lake ID
dat$lake.id <- as.numeric(as.factor(as.numeric(as.factor(dat$lake_id))))
# Create year ID
dat$year.id <- as.numeric(as.factor(as.numeric(as.factor(dat$year))))
nyear <- length(unique(dat$year.id))

# Sort data by site
dat <- dat %>% 
  arrange(lake.id)

# Effort offset term (maybe multiply eff_seconds by the number_of_netters?)
dat <- dat %>% 
  mutate(offsets = log(total_effort_1))
```

Lake-level predictors
```{r}
# Prepare lake-level predictors
lake_area <- summaryBy(log_acres ~ lake.id, data=dat, FUN=mean)
gdd <- summaryBy(gdd_wtr_5c ~ lake.id, data=dat, FUN=mean)

# Transform as needed and scale
lake_area2 <- as.numeric(scale(log(lake_area$log_acres.mean)))
gdd2 <- as.numeric(scale(gdd$gdd_wtr_5c.mean))
```

Observation-level predictors
```{r}
# Observation-level predictors, transform as needed and scale
annual_gdd <- as.numeric(scale(dat$anomaly_gdd_5)) # GDD up to day of sampling
post <- dat$post
x.treatment <- dat$treatment
stocked <- dat$STOCKED
```

Quick test with glmer
```{r}
m0 <-glmer(catch~ 1 + (1 | lake.id), data = dat, family = poisson(link = "log"), nAGQ = 100)
summary(m0)
# # 
m1 <- glmer(catch ~ post + treatment + post*treatment + STOCKED + (1|lake.id), nAGQ = 100,
             data=dat, family = poisson(link="log"))
summary(m1)
performance::check_zeroinflation(m1)

```

Check that data is in correct form
```{r}
n_post <- max(post)
n_trt <- max(x.treatment)
n_stocked <- max(stocked)
n_lake <- max(dat$lake.id)
n_year <- max(dat$year.id)


```

Load data; x = x; (dim(x)[2] + dim(x_obs)[2] + 1)
```{r}
data <- list(y = dat$catch, lake = dat$lake.id, N = dim(dat)[1], 
             n_lake = n_lake, offsets=dat$offsets, 
             year = dat$year.id, n_year = n_year, n_post = n_post, n_trt = n_trt, n_stocked = n_stocked,
             post = post, trt = x.treatment, stocked = stocked, survey_gdd = annual_gdd, 
             lake_area = lake_area2, gdd = gdd2)
```


```{r}
out <- stan(file='./pois_model_zip3_MRVrerun.stan', data = data,
            iter = 50000, warmup = 20000, chains=4, thin=1, cores=4, refresh = 1, control=list(adapt_delta=0.99, max_treedepth = 20))
print(out)

# Save all the stan output
#
saveRDS(out, file="./models/model_out_zip_CRrerun_nosecchi.rds")
#out <- readRDS(file='./models/model_out_zip_CRrerun_nosecchi.rds')
sum.stan <- summary(out)$summary %>% as.data.frame()

mcmc_trace(out, pars = c("sigma_year", "sigma_lake", "theta", 
                         "b_0", "b_post", "b_trt", "b_stocked", "b_post_trt", 
                         "b_survey_gdd", "b_lake_area", "b_gdd"))
```



```{r}
library(brms)
library(ggokabeito) 
options(scipen = 999)

data_ef <- dat %>% 
  dplyr::select(lake.id, year.id, offsets, catch, post, treatment, STOCKED, 
         julian_day_survey, gdd_wtr_5c, log_acres, median_gdd_5) %>%
  mutate(across(julian_day_survey:median_gdd_5, .names = "{col}_s", ~c(scale(.)))) %>%
  mutate(treatment = as.factor(treatment), post = as.factor(post), STOCKED = as.factor(STOCKED))
mean(data_ef$offsets)

prior <- c(set_prior("normal(0, 2.5)", class = "b"),
           set_prior("beta(2, 2)", class = "zi"),
           set_prior("normal(0, 5)", class = "Intercept"))

model.ef <- brm(catch ~ offset(offsets) + treatment*post + STOCKED + 
                  log_acres_s + gdd_wtr_5c_s + median_gdd_5_s + 
                  (1|year.id) + (1|lake.id), 
                family =  zero_inflated_poisson,
                data = data_ef, control = list(adapt_delta = 0.95),
                prior = prior,
                chains = 4, iter = 5000, warmup = 1500,
                cores = 4, seed = 1024)

saveRDS(model.ef, file="./models/model_out_brms.rds")
model.ef <- read_rds(file="./models/model_out_brms.rds")
summary(model.ef)
conditional_effects(model.ef)
sum.stan
  
invasion_epred <- model.ef %>%
  emmeans( ~ post*treatment, epred = T, re_formula = NA, type = "response") %>%
  gather_emmeans_draws() %>%
  median_qi(.width = c(.66, .95)) %>%
  mutate(across(where(is.numeric), ~ round(.x, 4))) %>%
  mutate(BACI = case_when(post == "0" & treatment == "1" ~ "1",
                          post == "1" & treatment == "1" ~ "2",
                          post == "0" & treatment == "0" ~ "3",
                          post == "1" & treatment == "0" ~ "4",
                          TRUE ~ NA))

new.data <- data.frame(post = c(0,1,0,1), treatment = c(1,1,0,0), post_trt = c(0,1,0,0), 
                       log_acres_s = c(0), gdd_wtr_5c_s = c(0), median_gdd_5_s = c(0),
                       julian_day_survey_s = c(0), STOCKED = c(0), offsets = 0.1966655)

model.ef %>% posterior_epred(newdata = new.data, allow_new_levels = T, sample_new_levels = "gaussian") %>% summarise_draws()

invasion_epred %>% 
  pivot_wider(names_from = .width, 
              values_from = c(".lower", ".upper", ".value")) %>%
  dplyr::select(!c(".interval", ".point", ".value_0.66")) %>%
  mutate(BACI = as.factor(BACI)) %>%
  ggplot(aes(x = BACI, y = .value_0.95, ymin = .lower_0.95, ymax = .upper_0.95))+
  geom_pointrange(aes(color = BACI), size = 1.5) + 
  geom_linerange(aes(x = BACI, ymin = .lower_0.66, ymax = .upper_0.66, color = BACI), size = 4)+
  theme_bw()+
  theme(legend.position = "none") +
  scale_x_discrete(limit = c("1", "2", "3", "4"),
                   labels = c("Pre ZM", "Post ZM", "Uninvaded Pre", "Uninvaded Post")) +
  theme(axis.text.x = element_text(color = "black", size = 14),
        axis.title.x = element_blank()) +
  labs(y = "Mean Predicted Walleye Catch per Unit Effort")
  
```

