---
title: "EF_WAE_ZM_Model1"
author: "Holly Kundel"
date: "11/15/2021"
output: html_document
---

This model is run using `ZM_LAKE` as zebra mussel invasion variable
- Only considered invaded if at least one EF survey was done the year of invasion or later
- No invasions after 2019 count (because no data post 2019)
- "0" indicates a lake is uninvaded (based on above criteria)
- "1" indicates a lake is invaded (based on above criteria)

Load in Required Packages
```{r}
#remove.packages(c("StanHeaders", "rstan", "brms"))
#install.packages("rstan", repos = "https://cloud.r-project.org/", dependencies = TRUE)
#install.packages("brms", repos = "https://cloud.r-project.org/", dependencies = TRUE)
library("rstan") # observe startup messages
library(brms)
library(sf)
library(RANN)
library(lme4)
library(tidyverse)
library(mnsentinellakes)
options(scipen = 999)
```

Load in Data
```{r, message=FALSE}
dat <- read_csv("./cleaned_data/yoy_catch_data.csv") %>% 
  mutate(survey_id = row_number(), 
         STOCKED = ifelse(FRY > 0,1,0))

nrow(dat[dat$catch==0,])/nrow(dat)


```
# Nearest Neighbor Code

- Assigning uninvaded lakes a "pseudo-invasion year" for BACI set up

Data setup
```{r, eval = F}
# Format dates
dat <- EF_WAE_ZM_data %>% 
  mutate(SAMPLING_START_DATE = mdy(SAMPLING_START_DATE),
         day = day(SAMPLING_START_DATE),
         month = month(SAMPLING_START_DATE),
         year = year(SAMPLING_START_DATE),
  )
```

Create Spatial sf Object
```{r}
# Create spatial sf object (4326 = WGS84)
dat.geo <- st_as_sf(dat, coords = c("x", "y"), crs = 4326)
```

Separate Invaded Lakes and Uninvaded Lakes

- Using column `ZM_LAKE` for this
- Only considered invaded if at least one EF survey was done the year of invasion or later
- No invasions after 2019 count (because no data post 2019)
- "0" indicates a lake is uninvaded (based on above criteria)
- "1" indicates a lake is invaded (based on above criteria)

```{r}
dat.geo.uninvaded <- dat.geo %>% 
  filter(year_confirmed == 0)
dim(dat.geo.uninvaded) #1116 surveys

dat.geo.invaded <- dat.geo %>% 
  filter(year_confirmed > 0)
dim(dat.geo.invaded) #322 surveys, adds up to correct total of 1438 surveys total
```

Get coordinate matrices
```{r}
uninvaded_coords <- do.call(rbind, st_geometry(dat.geo.uninvaded))
invaded_coords <- do.call(rbind, st_geometry(dat.geo.invaded))
invaded_coords <- cbind(invaded_coords, 1:nrow(invaded_coords)) # Index for identifying nearest lake

```

Find nearest neighbor
```{r}
closest <- nn2(invaded_coords[,1:2], uninvaded_coords,  k = 1, searchtype = "standard")
#str(closest)
#head(closest)
closest <- sapply(closest, cbind) 
```

Combine uninvaded data with nearest neighbor assignments
```{r}
dat.geo.uninvaded <- cbind(dat.geo.uninvaded, closest)
```

Give invaded lakes an identifier and select columns
```{r}
dat.geo.invaded$nn.idx <- 1:nrow(dat.geo.invaded)
dat.geo.invaded <- dat.geo.invaded %>% 
  dplyr::select(year_confirmed, nn.idx)
str(dat.geo.invaded)
head(dat.geo.invaded)

#str(dat.geo.uninvaded)
#str(dat.geo.invaded)
```

Merge uninvaded with invaded `YEAR_INFESTED` for use as "after" year in BACI (Before After Control Impact)
```{r}
dat.geo.invaded_merge <- st_set_geometry(dat.geo.invaded, NULL)

dat.geo.uninvaded.final <- dat.geo.uninvaded %>%
    left_join(dat.geo.invaded_merge, by = c("nn.idx"))
#str(dat.geo.uninvaded.final)
#summary(dat.geo.uninvaded.final)
# Bring back into full dataset
# Convert to dataframe
dat.geo.uninvaded.final.df <- st_set_geometry(dat.geo.uninvaded.final, NULL)
#class(dat.geo.uninvaded.final.df)
#str(dat.geo.uninvaded.final.df)

# Join dat.geo.uninvaded.final.df - want YEAR_INFESTED.y - with original data, dat
# Grab columns of interest
dat.geo.uninvaded.final.df <- dat.geo.uninvaded.final.df %>% 
  dplyr::select(survey_id, year_confirmed.y)
#dim(dat.geo.uninvaded.final.df)
#head(dat.geo.uninvaded.final.df)
#dim(dat)

# Final merge
final.dat3 <- dat %>%
    left_join(dat.geo.uninvaded.final.df, by = c("survey_id"))
#dim(final.dat3)
#head(final.dat3)
```

# Prepare to run model

Create BACI design variables
```{r}
# "Treatment" is lake-level invasion status, ZM (1 = ZM; 0 otherwise)
# Before/After time variable ("post" = 1 if sampled after invasion; 0 if before)
# Below, "year" is year of sampling and time is year infested
dat <- final.dat3 %>% 
  mutate(ZM_LAKE = ifelse(year_confirmed == 0, 0, 1),
         offsets = log(total_effort_1),
         time = ifelse(year_confirmed == 0, year_confirmed.y, year_confirmed),
         post = ifelse(year > time, 1, 0),
         treatment = ZM_LAKE # USING ZM_LAKE_HERE
         )%>%
  mutate(INVASION_STATUS_NEW = case_when(ZM_LAKE == 0 ~ "Uninvaded",
                                         ZM_LAKE == 1 & post == 0 ~ "PRE_ZM",
                                         ZM_LAKE == 1 & post == 1 ~ "POST_ZM"))

#head(dat)
#summary(dat)

#check data
dat %>%
  group_by(ZM_LAKE) %>%  
  count(post, ZM_LAKE) %>%
  ungroup() %>% 
  mutate(prop = prop.table(n))

###new addition of data filtering#######
#ensure at least one year pre and post
test = dat %>% 
  group_by(lake_id, post, ZM_LAKE) %>% 
  summarise(count=n()) %>% 
  filter(count > 0) %>% ungroup() %>% 
  group_by(lake_id, ZM_LAKE) %>% 
  summarise(pre.post.test = n()) %>% 
  filter(pre.post.test==2)


dat <- dat %>% filter(lake_id %in% test$lake_id)
```

# Number of lakes stocked
```{r}
invasions <- dat %>% select(lake_id, year_confirmed)

stocking <- read.csv("./data/mn_wae_stocking.csv") %>% 
  rename(lake_id = DOW, year = Year) %>%
  mutate(lake_id = ifelse(lake_id == "04003500", "04003501", lake_id),
         lake_id = fixlakeid(lake_id)) %>%
  select(lake_id, year, FRY, fry.pa) %>% 
  filter(lake_id %in% dat$lake_id) %>% 
  merge(invasions, by = "lake_id") %>%
  mutate(invaded = ifelse(year_confirmed > year, 1, 0))
  
```


Quick test with glmer
```{r, eval = F}
m0 <-glmer(catch~ 1 + (1 | lake.id), data = dat, family = poisson(link = "log"), nAGQ = 100)
summary(m0)
# # 
m1 <- glmer(catch ~ post + treatment + post*treatment + STOCKED + (1|lake.id), 
               family = poisson(link = "log") nAGQ = 100, data=dat)
summary(m1)
performance::check_zeroinflation(m1)

```

# Models
```{r}
to_keep <- c("anomaly_gdd_5_s","anomaly_gdd_5", "STOCKED", "median_gdd_5_s", "median_gdd_5", "lake_area_s", "log_acres")

dat_standardized <- dat %>% 
  mutate(anomaly_gdd_5_s = as.numeric(scale(anomaly_gdd_5)),
         lake_area_s = as.numeric(scale(log_acres)),
         median_gdd_5_s = as.numeric(scale(median_gdd_5)),
         #lake_id = as.factor(lake_id),
         year = as.factor(year)) %>%
  dplyr::select(lake_id, year, catch, offsets, total_effort_1, post, treatment, x, y, matches(to_keep))

write.csv(dat_standardized, "./cleaned_data/ef_for_model.csv", row.names = F)
dat_standardized <- read.csv("./cleaned_data/ef_for_model.csv")

temp <- dat_standardized %>% group_by(lake_id) %>% count()
median(temp$n)

length(unique(temp$lake_id))
min(dat_standardized$year); max(dat_standardized$year)
```


```{r}
priors = c(prior(normal(0, 5), class = "Intercept"),
           prior(normal(0, 5), class = "b"))

# Zero inflated poisson with covariates
zip_ef_model <- 
  brm(catch ~ offset(offsets) + 
      post + treatment + post:treatment + STOCKED + anomaly_gdd_5_s + median_gdd_5_s + 
      lake_area_s +
      (1 | year) +              
      (1 | lake_id),
    family = zero_inflated_poisson(), data = dat_standardized, prior = priors,
    control = list(max_treedepth = 25),
    chains = 4, iter = 10000, warmup = 2500, thin = 2, seed = 1024, save_pars=save_pars(all=TRUE),
    file = "./models/zip_ef_model_brms.rds")
#summary(zip_ef_model)
#plot(zip_ef_model)

# Zero inflated poisson NO covariates
zip_ef_model_nocovs <- 
  brm(catch ~ offset(offsets) + 
      post + treatment + post:treatment +
      (1 | year) +              
      (1 | lake_id), 
    family = zero_inflated_poisson(), data = dat_standardized, prior = priors,
    control = list(max_treedepth = 25),
    chains = 4, iter = 10000, warmup = 2500, thin = 2, seed = 1024, save_pars=save_pars(all=TRUE),
    file = "./models/zip_ef_model_nocovs_brms.rds")
#summary(zip_ef_model_nocovs)

# Negative Binomial with covariates
nb_ef_model <- 
  brm(catch ~ offset(offsets) + 
      post + treatment + post:treatment + STOCKED + anomaly_gdd_5_s + median_gdd_5_s + 
      lake_area_s + 
      (1 | year) +              
      (1 | lake_id), 
    family = negbinomial(), data = dat_standardized, prior = priors,
    control = list(max_treedepth = 25),
    chains = 4, iter = 10000, warmup = 2500, thin = 2, seed = 1024, save_pars=save_pars(all=TRUE),
    file = "./models/nb_ef_model_brms.rds")
#plot(nb_ef_model)
#summary(nb_ef_model)


# Negative Binomial NO covariates
nb_ef_model_nocovs <- 
  brm(catch ~ offset(offsets) + 
      post + treatment + post:treatment +
      (1 | year) +              
      (1 | lake_id), 
    family = negbinomial(), data = dat_standardized, prior = priors,
    control = list(max_treedepth = 25),
    chains = 4, iter = 10000, warmup = 2500, thin = 2, seed = 1024, save_pars=save_pars(all=TRUE), 
    file = "./models/nb_ef_model_nocovs_brms.rds")
#summary(nb_ef_model_nocovs)

# Zero inflated Negative Binomial with covariates
zipnb_ef_model <- 
  brm(catch ~ offset(offsets) + 
      post + treatment + post:treatment + STOCKED + anomaly_gdd_5_s + median_gdd_5_s + 
      lake_area_s +
      (1 | year) +              
      (1 | lake_id), 
    family = zero_inflated_negbinomial(), data = dat_standardized, prior = priors,
    control = list(max_treedepth = 25),
    chains = 4, iter = 10000, warmup = 2500, thin = 2, seed = 1024, save_pars=save_pars(all=TRUE),
    file = "./models/zipnb_ef_model_brms.rds")
#plot(zipnb_ef_model)
#summary(zipnb_ef_model)
```


loo - fits model with all data minus one then calculates the probability of that observation (0-1; Predictive density). It then turns that into a log-density (log predictive density). It then does this for every observation and sums the log predictive density (expected log predictive density; elpd_loo). Higher elpd_loo is BETTER (since elpd_loo is negative, closer to zero is better) Log predictive density is the calculation if you never leave out data (lpd). p_loo is lpd - elpd_loo and should always be less than zero ( I think?)

# Model selection, and other fun tidbits
```{r}
output_nb <- loo(nb_ef_model, moment_match = TRUE)
output_nb_nocovs <- loo(nb_ef_model_nocovs, moment_match = TRUE)
output_zipnb <- loo(zipnb_ef_model, moment_match = TRUE)

# These take quite a bit to run and dont work (high pareto K value)
output_zip <- loo(zip_ef_model, moment_match = TRUE)
output_zip_nocovs <- loo(zip_ef_model_nocovs, moment_match = TRUE)

loo_compare(output_nb, output_zipnb, output_nb_nocovs)

start <- Sys.time()
kfoldzip <- kfold(zip_ef_model, k = 10) # -59079.2	SE = 4751.4	
#kfoldnb <- kfold(nb_ef_model, k = 10) # elpd_kfold	-6989.2	SE 72.7
#kfoldzionb <- kfold(zipnb_ef_model, k = 10) # elpd_kfold	-6976.7	SE 72.6	
#print(kfoldzip); print(kfoldnb); print(kfoldzionb)
kfold_results <- list(zip = kfoldzip, nb = kfoldnb, zinb = kfoldzionb)
saveRDS(kfold_results, "./models/kfold_results.rds")
kfold_results <- readRDS("./models/kfold_results.rds")
end <- Sys.time()
total <- end - start

kfold_results$zip
kfold_results$nb
kfold_results$zinb


zipnb_ef_model <- add_criterion(zipnb_ef_model, "loo")
nb_ef_model <- add_criterion(nb_ef_model, "loo")
zip_ef_model <- add_criterion(zip_ef_model, "loo")

loo_compare(zipnb_ef_model, nb_ef_model, zip_ef_model)


summary(zip_ef_model)         
#post:treatment = -0.04
summary(nb_ef_model)          
#post:treatment = -0.54
summary(zip_ef_model_nocovs)
#post:treatment =
summary(nb_ef_model_nocovs)
#post:treatment = -0.56
summary(zipnb_ef_model)
#post:treatment = -0.45

bayes_R2(nb_ef_model)
pp_check(nb_ef_model, ndraws = 100)
pp_check(nb_ef_model, type='stat', stat='mean')
pp_check(nb_ef_model, type='error_scatter_avg')

bayes_R2(zipnb_ef_model)
pp_check(zipnb_ef_model, ndraws = 100)
pp_check(zipnb_ef_model, type='stat', stat='mean')
pp_check(zipnb_ef_model, type='error_scatter_avg')
```

