---
title: "Figures"
author: "Holly Kundel"
date: "`r Sys.Date()`"
output: html_document
---


Load in required packages

```{r, message = FALSE, warning = FALSE}
library(rstan)
library(bayesplot)
library(dplyr)
library(lubridate)
library(tidyr)
library(stringr)
library(ggplot2)
library(forcats)
library(ggmcmc)
library(ggthemes)
library(readr)
library(readxl)
library(doBy)
library(RColorBrewer)
library(tibble)
library(tidybayes)
library(viridis)
library(ggspatial)
library(maps)
library(cowplot)
# library(ggsn)
library(grid)
library(data.table)
library(shinystan)
```

# Data


```{r, message=FALSE}
GN_Final <- read_csv("./cleaned_data/gn_model_data.csv") %>%
  mutate(ZM_LAKE = ifelse(year_confirmed == 0, "Uninvaded", "ZM Invaded"))

#out <- readRDS(file='./models/model_out_zip_GN_nosecchi.rds')
out <- readRDS(file="./models/model_out_zip_GN_nosecchi_3gddanomalyCR.rds" )


# Convert mcmc draws to matrix
posterior <- as.matrix(out)




posterior_df <- as.data.table(posterior)[ ,c(1:16),] #take only the covariate columns

```




# Figure 1

Map showing locations of lakes and whether or not they have zebra mussels as of 2019

```{r Figure 1, warning = FALSE, message = FALSE}
MN <- map_data("state")%>% filter(region=="minnesota")
MN_Counties <- map_data("county")%>%
  filter(region == "minnesota")

# walleye lakes, ZM in red, rest in blue
GN_Lakes_ZM <- ggplot(data = MN_Counties, mapping = aes(x = long, y = lat, group=group))+
  coord_fixed(1.3) +
  geom_polygon(color= "gray85", fill = "gray90")+
  geom_point(data=GN_Final, aes(x= x, y = y, group=25, color=as.factor(ZM_LAKE),
                                      shape = as.factor(ZM_LAKE)),alpha = 0.3, size =3) +
  scale_color_manual(name = "ZM Invasion Status", labels = c("Uninvaded", "ZM Invaded"), values = c("#0571b0", "#ca0020"))+
  scale_shape_manual(name = "ZM Invasion Status", labels = c("Uninvaded", "ZM Invaded"), values = c(16, 15))+
  labs(color = "ZM Invasion Status",x= "Longitude", y ="Latitude")+
  theme_void()+
  theme(legend.position = "bottom",
        legend.box.background = element_rect(color = "gray85"),
        legend.box.margin = margin(t = 2, r = 4, b = 2, l = 4),
        legend.key = element_rect(color = "gray90"))+
  coord_sf(crs = 26991) #EPSG MN North https://epsg.io/26991

GN_Lakes_ZM
```

Bounding box and North America Map for inset
```{r country}
canadamap <- map_data("world", c("Canada")) 

mexicomap <- map_data("world", c("Mexico"))

all_states <- map_data("state")

lakes <- map_data("lakes") #pulls in great lakes and large lakes

states <- subset(all_states, region %in% c("minnesota"))

xlim <- range(states$long) + c(-.3, 0.1) 
ylim <- range(states$lat) + c(-0.7, .65)
# the +/- 0.3 creates an offset so the bounding box is not counting off MN

N_AM_map <- ggplot() +  
  geom_map(data = canadamap, aes(map_id = region, group = subregion),  
           map = canadamap, color = "black", fill = "grey90") +
  # creates a map of Canada
  geom_map(data = mexicomap, aes(map_id = region, group = subregion),  
           map = mexicomap, color = "black", fill = "grey90")+
  #creates a map of Mexico
  geom_map(data = all_states, aes(map_id = region), 
           map = all_states, color = "black", fill = "grey90") +
  # Creates a map of US states with the existing Canada map
  geom_map(data = lakes,aes(map_id = region, group = subregion),
           map = lakes, color = "black", fill = "lightskyblue2")+
  #makes the lines around the great lakes cleaner, and can fill them with blue
  geom_rect(aes(xmin = xlim[1], xmax = xlim[2], ymin = ylim[1], ymax = ylim[2]), 
            colour = "red", fill = NA, lwd = 1) + 
  # creates a red rectangle around the y- and x-lim
  coord_map(xlim = c(-125, -66), ylim = c(25, 50)) +
  # defines the cutoff/edges of the map
  labs(x = NULL, y = NULL) + theme_bw() + theme(panel.grid = element_blank())+
  theme(panel.background = element_rect(fill = "lightskyblue2", color = "black"),
        axis.text.x = element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank()) #remove y axis ticks
N_AM_map
```

Inset maps
```{r inset}
Final_Inset_Map <- ggdraw()+
  draw_plot(GN_Lakes_ZM +
              # ggsn::scalebar(data = MN_Counties, dist = 50, dist_unit = "km", st.size = 2, 
              #           transform = TRUE, model = 'WGS84', height = 0.01, location = "bottomright",
              #           box.color = c("black", "grey"))+
              annotation_north_arrow(location = "br", 
                         which_north = "grid",
                         pad_x = unit(0.7, "in"),
                         pad_y = unit(0.7, "in"),
                         style = north_arrow_orienteering,
                         width = unit(.7, "cm"),
                         height = unit(.7, "cm")))+ #adds scale bar
  draw_plot(N_AM_map, x = 0.7, y = 0.3, width = .2, height = .15)

Final_Inset_Map
#
ggsave("./figures/figure_1_GN.jpeg", height = 10, width = 9)
```

# Figure 2

BACI Figure

Data
```{r}
NoZM_Pre <- exp(posterior[,"b_0"]) 
mean(NoZM_Pre)
quantile(NoZM_Pre, c(0.025, 0.975))
# Control Post
NoZM_Post <- exp(posterior[,"b_0"] + posterior[,"b_post"]) 
mean(NoZM_Post)
quantile(NoZM_Post, c(0.025, 0.975))
# ZM Treatment Pre
ZM_Pre <- exp(posterior[,"b_0"] + posterior[,"b_trt"]) 
mean(ZM_Pre)
quantile(ZM_Pre, c(0.025, 0.975))
# ZM Treatment Post
ZM_Post <- exp(posterior[,"b_0"] + posterior[,"b_post"] +  posterior[,"b_trt"] +  posterior[,"b_post_trt"]) 
mean(ZM_Post)
quantile(ZM_Post, c(0.025, 0.975))

BACI_Estimates1 <- tibble(
  Category = c("No_ZM_Pre", "No_ZM_Post", "ZM_Pre", "ZM_Post"),
  Mean = c(mean(NoZM_Pre), mean(NoZM_Post), mean(ZM_Pre), mean(ZM_Post)),
  Quant_2.5 = c(quantile(NoZM_Pre, 0.025), quantile(NoZM_Post, 0.025), quantile(ZM_Pre, 0.025), quantile(ZM_Post, 0.025)),
  Quant_97.5 = c(quantile(NoZM_Pre, 0.975), quantile(NoZM_Post, 0.975), quantile(ZM_Pre, 0.975), quantile(ZM_Post, 0.975)),
  ZM = c("Uninvaded", "Uninvaded", "ZM", "ZM"),
  Quant_83 = c(quantile(NoZM_Pre, 0.83), quantile(NoZM_Post, 0.83), quantile(ZM_Pre, 0.83), quantile(ZM_Post, 0.83)),
  Quant_17 = c(quantile(NoZM_Pre, 0.17), quantile(NoZM_Post, 0.17), quantile(ZM_Pre, 0.17), quantile(ZM_Post, 0.17)))

BACI_PLOT_DAT <- BACI_Estimates1 %>%
  mutate(Levels = fct_relevel(Category, "No_ZM_Pre", "No_ZM_Post", "ZM_Pre", "ZM_Post"))
```


```{r, warning = FALSE}
brewer.pal(n = 2, "Set1") #E41A1C = red, #377EB8 = blue

BACI_Figure <-ggplot(BACI_PLOT_DAT , aes(x = Levels, y = Mean, ymin = Quant_2.5, ymax = Quant_97.5))+
  geom_pointrange(aes(color = ZM),size = 1.5)+
  geom_linerange(aes(x = Levels, ymin = Quant_17, ymax = Quant_83, color = ZM), size = 4)+
  #geom_segment(aes(x = "No_ZM_Pre", y = 4.189089, xend = "No_ZM_Post", yend = 4.464494), color = "#377EB8", linetype = "dashed")+
  #geom_segment(aes(x = "ZM_Pre", y = 9.186239, xend = "ZM_Post", yend = 5.449405), color ="#E41A1C", linetype = "dashed")+
  scale_color_brewer(palette = "Set1", direction = -1)+
  labs(y = "Mean Predicted Walleye Catch per Unit Effort")+
  theme_bw()+
  theme(legend.position = "none")+
  scale_x_discrete(limit = c("ZM_Pre", "ZM_Post", "No_ZM_Pre", "No_ZM_Post"),
                     labels = c("Pre ZM", "Post ZM", "Uninvaded Pre", "Uninvaded Post"))+
  theme(axis.text.x = element_text(family = "Arial", color = "black", size = 14),
        axis.title.x = element_blank())#,
        
BACI_Figure
#ggsave("./figures/figure_2_GN.jpeg", height = 10, width = 10)
```


# Figure 3
Posterior Distributions

```{r}
color_scheme_set(scheme = "blue")
post_dist_blue2 <- mcmc_areas(
    posterior,
    pars = c("b_trt", "b_post", "b_post_trt","b_lake_area", "b_fry", "b_fgl", 
             "b_survey_gdd_sum", "b_gdd"),
    prob = 0.95, # 95% intervals
    prob_outer = 0.99, # 99%
    point_est = "median",
    area_method = "scaled height"
  ) + 
  geom_vline(xintercept = 0, color = "red", linetype = "dashed") + 
  scale_y_discrete(limit = c("b_trt", "b_post", "b_post_trt","b_lake_area", "b_fry", "b_fgl", 
                             "b_survey_gdd_sum", "b_gdd"),
                   labels = c("Zebra Mussels", "Post Invasion", "ZM * Post", "Lake Area", 
                              "Stocked (fry)", "Stocked (fingerlings)", 
                              "GDD (age 0 - 2)", "Lake median GDD")) +
  theme(axis.text.x = element_text(color = "black"),
        axis.text.y = element_text(color = "black", size = 14, family = "sans"))+
  theme_classic()

post_dist_blue2
#ggsave("./figures/figure_3_GN.jpeg", height = 10, width = 10)
```


# Figure 4

Stocking Figure

```{r, warning = FALSE}
# get stocking estimates

not_stocked <- exp(posterior[,"b_0"]) 

stocked_fry <- exp(posterior[,"b_0"] + posterior[,"b_fry"]) 

stocked_fgl <- exp(posterior[,"b_0"] + posterior[,"b_fgl"])


stocking_estimates <- tibble(
  Stocking_Status = c("Not Stocked", "Fry Stocked", "Fingerling Stocked"),
  Mean = c(mean(not_stocked), mean(stocked_fry), mean(stocked_fgl)),
  Quant_2.5 = c(quantile(not_stocked, 0.025), quantile(stocked_fry, 0.025), quantile(stocked_fgl, 0.025)),
  Quant_97.5 = c(quantile(not_stocked, 0.975), quantile(stocked_fry, 0.975), quantile(stocked_fgl, 0.975)),
  Quant_83 = c(quantile(not_stocked, 0.83), quantile(stocked_fry, 0.83), quantile(stocked_fgl, 0.83)),
  Quant_17 = c(quantile(not_stocked, 0.17), quantile(stocked_fry, 0.17), quantile(stocked_fgl, 0.17)))


stocking_estimates_DAT <- stocking_estimates %>%
  mutate(Levels = fct_relevel(Stocking_Status, "Not Stocked", "Fry Stocked", "Fingerling Stocked"))


#get hex codes for colors
brewer.pal(n = 4, "Set1") #E41A1C = red, #377EB8 = blue, #4DAF4A = green, #984EA3 = purple

Stocking_Figure <-ggplot(stocking_estimates_DAT , aes(x = Stocking_Status, y = Mean, ymin = Quant_2.5, ymax = Quant_97.5))+
  geom_pointrange(aes(color = Stocking_Status), size = 1.5)+
  geom_linerange(aes(x = Levels, ymin = Quant_17, ymax = Quant_83, color = Stocking_Status), size = 4)+
  scale_discrete_manual(aesthetics = c("color", "fill"), values = c("#4DAF4A", "#984EA3", "#377EB8"))+
  labs(y = "Mean Predicted Walleye Catch per Unit Effort")+
  theme_bw()+
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(family = "Arial", color = "black", size = 14),
        axis.title.x = element_blank())
        
Stocking_Figure
#ggsave("./figures/figure_4_GN.jpeg", height = 10, width = 10)
```


# Figure 5

Environmental effects

Model Data
```{r}
dat <- read.csv('./cleaned_data/gn_model_data.csv') %>%
  mutate(ZM_LAKE = ifelse(year_confirmed == 0, 0, 1))  %>% 
  mutate(time = ifelse(year_confirmed == 0, year_confirmed.y, year_confirmed),
         post = ifelse(year > time, 1, 0),
         treatment = ZM_LAKE # USING ZM_LAKE_HERE
         )%>%
  mutate(INVASION_STATUS_NEW = case_when(ZM_LAKE == 0 ~ "Uninvaded",
                                         ZM_LAKE == 1 & post == 0 ~ "PRE_ZM",
                                         ZM_LAKE == 1 & post == 1 ~ "POST_ZM"),
         log_acres = log(acres))


# Create lake ID
dat$lake.id <- as.numeric(as.factor(as.numeric(as.factor(dat$lake.id))))
# Create year ID
dat$year.id <- as.numeric(as.factor(as.numeric(as.factor(dat$year))))
nyear <- length(unique(dat$year.id))

# Sort data by site
dat <- dat %>% 
  arrange(lake.id)

```

Lake-level predictors
```{r}
# Prepare lake-level predictors
lake_area <- summaryBy(log_acres ~ lake.id, data=dat, FUN=mean)
gdd <- summaryBy(median_gdd_5 ~ lake.id, data=dat, FUN=mean)
#secchi <- summaryBy(median_secchi_lake ~ lake.id, data=dat, FUN=mean) # not used in model

# Transform as needed and scale
lake_area2 <- as.numeric(scale(lake_area$log_acres.mean))
gdd2 <- as.numeric(scale(gdd$median_gdd_5.mean))
#secchi2 <- as.numeric(scale(secchi$median_secchi_lake.mean)) # not used in model

```

Observation-level predictors
```{r}
gdd_sum <- as.numeric(scale(dat$gdd_wtr_5c_year2 + dat$gdd_wtr_5c_year1 + dat$gdd_wtr_5c_year0))
```



GDD annual 5C from year of survey (sum)
```{r}
# Begin predict code for first three year GDD

b_gdd_sum <- posterior_df$b_survey_gdd_sum
b_0 <- posterior_df$b_0

sum <- (dat$gdd_wtr_5c_year0 + dat$gdd_wtr_5c_year1 + dat$gdd_wtr_5c_year2) - (3*dat$median_gdd_5) 

lower_gdd = min(sum)
upper_gdd = max(sum)
gdd_range <- seq(lower_gdd, upper_gdd, by= 5) 

gdd_range_standardized <- wiqid::standardize2match(gdd_range, sum)

# Output storage DF
gdd.plot.anomaly <- data.frame(gdd = gdd_range, 
                        gdd_s = gdd_range_standardized, 
                        mean = numeric(length(gdd_range)), 
                        low = numeric(length(gdd_range)), high = numeric(length(gdd_range)))

for (i in 1:nrow(gdd.plot.anomaly)){
  pred.gdd = exp(b_0 + (b_gdd_sum * gdd.plot.anomaly[i, 2]))
  gdd.plot.anomaly[i, 3] = mean(pred.gdd)
  gdd.plot.anomaly[i, 4:5] = quantile(pred.gdd, c(.025, .975))
}

# Plot
gdd_Plot_sum <- gdd.plot.anomaly %>% 
  ggplot(aes(x = gdd)) +
  geom_line(aes(x= gdd, y = mean), size = 2)+
  geom_ribbon(aes(ymin = low, ymax = high), alpha= .2) +
  labs(x = "GDD Anomaly from Age-0, Age-1, Age-2 (base 5°C)", y = "Predicted CPUE")+
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) +
  theme_classic()


gdd_Plot_sum

```


Lake area
```{r}
# Begin predict code

b_area <- posterior_df$b_lake_area
b_0 <- posterior_df$b_0

# Lake area range
lower_area = min(dat$log_acres)
upper_area = max(dat$log_acres)
log_lake_area_range <- seq(lower_area, upper_area, by= 0.1) #935 numbers

log_lake_area_range_standardized <- wiqid::standardize2match(log_lake_area_range, dat$log_acres)

area.plot <- data.frame(area = log_lake_area_range, 
                        area_s = log_lake_area_range_standardized, 
                        mean = numeric(length(log_lake_area_range)), 
                        low = numeric(length(log_lake_area_range)), high = numeric(length(log_lake_area_range)))

for (i in 1:nrow(area.plot)){
  pred.area = exp(b_0 + (b_area *area.plot[i, 2]))
  area.plot[i, 3] = mean(pred.area)
  area.plot[i, 4:5] = quantile(pred.area, c(.025, .975))
}

area_Plot <- area.plot %>% 
  ggplot(aes(x = area))+
  geom_line(aes(x=area, y = mean), size = 2)+
  geom_ribbon(aes(ymin = low, ymax = high), alpha= .2)+
  labs(x = "log Lake Area (acres)", y = "Predicted CPUE")+
  theme_classic()

area_Plot
```


Median GDD (1993 - 2019), lake level
```{r, eval = F}
# Begin predict code
b_gdd_LL <- posterior_df$b_gdd
b_0 <- posterior_df$b_0

# Range of observed lake level GDD's (base 5)
range(dat$median_gdd_5)
lower_gdd = range(dat$median_gdd_5)[1]
upper_gdd = range(dat$median_gdd_5)[2]


gdd_LL_range <- seq(lower_gdd, upper_gdd, by= 5)
# Standardized the range
gdd_LL_range_standardized <- wiqid::standardize2match(gdd_LL_range, dat$median_gdd_5)

# Output storage DF
gdd.plot <- data.frame(gdd = gdd_LL_range, 
                        gdd_s = gdd_LL_range_standardized, 
                        mean = numeric(length(gdd_LL_range)), 
                        low = numeric(length(gdd_LL_range)), high = numeric(length(gdd_LL_range)))

for (i in 1:nrow(gdd.plot)){
  pred.gdd = exp(b_0 + (b_gdd_LL * gdd.plot[i, 2]))
  gdd.plot[i, 3] = mean(pred.gdd)
  gdd.plot[i, 4:5] = quantile(pred.gdd, c(.025, .975))
}

# Plot
gdd_LL_Plot <- gdd.plot %>% 
  ggplot(aes(x = gdd))+
  geom_line(aes(x= gdd, y = mean), size = 2)+
  geom_ribbon(aes(ymin = low, ymax = high), alpha= .2)+
  labs(x = "Median Growing Degree Days", y = "Predicted CPUE")+
  theme_classic()
gdd_LL_Plot
```



# Combine into one figure
```{r}
Fig_5 <- plot_grid(gdd_LL_Plot, gdd_Plot_sum, area_Plot,
          labels = c("A", "B", "C"),
          label_size = 14)
Fig_5
#ggsave("./figures/figure_5_GN.jpeg", height = 10, width = 10)
```





# Summaries referenced in the manuscript, Table 1, and figures


Additional Data Summaries

Calculate percent change in pre vs post - zm
```{r}
#calculate percent change in pre-post
zm_diff = ZM_Pre - ZM_Post
zm_diff_prop = zm_diff/ZM_Pre

mean(zm_diff) #3.736835 walleye per hour
mean(zm_diff_prop) #40.67%
quantile(zm_diff_prop, c(0.025, 0.975))

uninvaded_diff = NoZM_Pre - NoZM_Post
mean(uninvaded_diff) #-0.2754049
uninvaded_prop = uninvaded_diff/NoZM_Pre
mean(uninvaded_prop)
quantile(uninvaded_prop, c(0.025, 0.975)) #-0.06585174 or 6.59% increase post


```

Effect of Stocking, stocking led to higher catch rates
```{r}
# Not stocked
Not_stocked <- exp(posterior[,"b_0"]) 
mean(Not_stocked) # 4.189089
quantile(Not_stocked, c(0.025, 0.975)) 

#Stocked (fry)
stocked_fry <- exp(posterior[,"b_0"] + posterior[,"b_fry"]) 
mean(stocked_fry) 
quantile(stocked_fry, c(0.025, 0.975)) 

#Stocked (fry)
stocked_fgl <- exp(posterior[,"b_0"] + posterior[,"b_fgl"]) 
mean(stocked_fgl) 
quantile(stocked_fgl, c(0.025, 0.975)) 



### Stocking effect (fry) ###
stocking_diff= stocked_fry - Not_stocked
mean(stocking_diff) # 6.949217
quantile(stocking_diff, c(0.025, 0.975)) 

# percent change with stocking (fry)
stocking_prop = stocking_diff/Not_stocked
mean(stocking_prop) 
quantile(stocking_prop, c(0.025, 0.975)) 


### Stocking effect (fgl) ###
stocking_diff_fgl = stocked_fgl - Not_stocked
mean(stocking_diff_fgl) 
quantile(stocking_diff_fgl, c(0.025, 0.975)) 

# percent change with stocking (fgl)
stocking_prop_fgl = stocking_diff_fgl/Not_stocked
mean(stocking_prop_fgl) 
quantile(stocking_prop_fgl, c(0.025, 0.975)) 
```

Basic Summary Info (Table 1)
```{r}
final.dat3 <- read_csv("R and stan code/Data/NN_EF_WAE_ZM_Model_MRVrerun.csv")

#surveys done in invaded and uninvaded lakes
Model_3_Summary <- final.dat3 %>%
  mutate(time = ifelse(!is.na(YEAR_INFESTED), YEAR_INFESTED, YEAR_INFESTED.y),
         post = ifelse(year > time, 1, 0))%>%
  mutate(Category = case_when(ZM_LAKE == 0 ~ "Uninvaded",
                                         ZM_LAKE == 1 & post == 0 ~ "PRE_ZM",
                                         ZM_LAKE == 1 & post == 1 ~ "POST_ZM"))%>%
  group_by(Category)%>%
  summarise(Total_Surveys = n()) #99 surveys post ZM, 223 surveys pre ZM, 1116 surveys uninvaded

#number of surveys done in each category
Model_3_Summary2 <- final.dat3 %>%
  mutate(time = ifelse(!is.na(YEAR_INFESTED), YEAR_INFESTED, YEAR_INFESTED.y),
         post = ifelse(year > time, 1, 0))%>%
  mutate(Category = case_when(ZM_LAKE == 0 & post == 0 ~ "Uninvaded_PRE",
                              ZM_LAKE == 0 & post == 1 ~ "Uninvaded_POST",
                                         ZM_LAKE == 1 & post == 0 ~ "PRE_ZM",
                                         ZM_LAKE == 1 & post == 1 ~ "POST_ZM"))%>%
  group_by(Category)%>%
  summarise(Total_Surveys = n()) # 99 post ZM, 223 pre zm, 181 uninvaded post, 935 uninvaded pre


# values of survey level covariates
Model_3_Summary_Survey_Level <- final.dat3 %>%
  mutate(time = ifelse(!is.na(YEAR_INFESTED), YEAR_INFESTED, YEAR_INFESTED.y),
         post = ifelse(year > time, 1, 0))%>%
  mutate(Category = case_when(ZM_LAKE == 0 & post == 0 ~ "Uninvaded_PRE",
                              ZM_LAKE == 0 & post == 1 ~ "Uninvaded_POST",
                                         ZM_LAKE == 1 & post == 0 ~ "PRE_ZM",
                                         ZM_LAKE == 1 & post == 1 ~ "POST_ZM"))%>%
  dplyr::select(SURVEY_YEAR, DD5, SURFACE_TEMP_F, AREA_HECTARES, MAX_DEPTH_M ,MEDIAN_93_19_GDD, TOTAL_CATCH,CPUE, YEAR_INFESTED , Category) %>%
  group_by(Category)%>%
  summarise_all(.funs = c(MEAN = mean.default, MAX = max, MIN = min), na.rm = TRUE) # %>%
  # select("Category", contains("CPUE")) #<- No worky
  

#write_csv(Model_3_Summary_Survey_Level, "Survey_Level_Summaries.csv")
# 
#values of lake level co-variates
Model_3_Summary_Lake_Level <- final.dat3 %>%
  mutate(time = ifelse(!is.na(YEAR_INFESTED), YEAR_INFESTED, YEAR_INFESTED.y),
         post = ifelse(year > time, 1, 0))%>%
  dplyr::select(36,37,38,44, 54,56, 61:63)%>%
  group_by(ZM_LAKE)%>%
  summarise_all(.funs = c(MEAN = mean, MAX = max, MIN = min), na.rm = TRUE)

# number of lakes in each category
Model_3_Summary_Lake <- final.dat3 %>%
  mutate(time = ifelse(!is.na(YEAR_INFESTED), YEAR_INFESTED, YEAR_INFESTED.y),
         post = ifelse(year > time, 1, 0))%>%
  mutate(Category = case_when(ZM_LAKE == 0 ~ "Uninvaded",
                                         ZM_LAKE == 1 & post == 0 ~ "PRE_ZM",
                                         ZM_LAKE == 1 & post == 1 ~ "POST_ZM"))%>%
  group_by(Category, DOW_KITTLE)%>%
  summarise(Total_Surveys = n())%>%
  group_by(Category)%>%
  summarise(Total_Lake_by_cat = n())
# 282 uninvaded lakes, 33 Pre-ZM, 33 Post-ZM lakes, 348 lakes total

# years surveys were done in each category
Model_3_Summary_Survey_Year <- final.dat3 %>%
  mutate(time = ifelse(!is.na(YEAR_INFESTED), YEAR_INFESTED, YEAR_INFESTED.y),
         post = ifelse(year > time, 1, 0))%>%
  mutate(Category = case_when(ZM_LAKE == 0 & post == 0 ~ "Uninvaded_PRE",
                              ZM_LAKE == 0 & post == 1 ~ "Uninvaded_POST",
                                         ZM_LAKE == 1 & post == 0 ~ "PRE_ZM",
                                         ZM_LAKE == 1 & post == 1 ~ "POST_ZM"))%>%
  dplyr::select(Category, SURVEY_YEAR)%>%
  group_by(Category)%>%
  summarise_all(.funs = c(MEAN = mean, MAX = max, MIN = min), na.rm = TRUE)
```

Posterior probabilities
```{r}
#GDD Lake Level average b_gdd

GDD_LL <- posterior_df%>%
  dplyr::select("b_gdd")%>%
  filter(b_gdd > 0)
nrow(GDD_LL)/75000 #0.7629467

#GDD survey Level average b_survey_gdd

GDD_LL <- posterior_df%>%
  dplyr::select(b_survey_gdd)%>%
  filter(b_survey_gdd > 0)
nrow(GDD_LL)/75000 # 0.14944

```

Number of ZM Infested lakes vs number of ZM Infested walleye lakes
First do with the data to 2019 (used here) 
THEN compare to the data that is most up to date
```{r}
# Read in infested waters list
  # can be downloaded here: https://www.dnr.state.mn.us/invasives/ais/infested.html
iw <- read_excel("R and stan code/Data/infested-waters_03_21_2023.xlsx", skip = 1)
iw <- iw[, -dim(iw)[2]]

# Change column names

colnames(iw) <- c("waterbody_name", "county", "species", "year", "year_conf", "dowlknum")

iw$connected <- factor(ifelse(grepl("connect", iw$year_conf) | grepl("Connect", iw$year_conf) | grepl("conect", iw$year_conf),
                              "connected", "confirmed"))


# Fixing some dowlknum entries
# CC-LLLL-BB, where C=county, L=lake or parent dow, B=basin, is a nice convention to standardize to
unique(iw$dowlknum[!grepl("-", iw$dowlknum)])

iw$dowlknum[iw$dowlknum=="NA"] <- NA
iw$dowlknum[iw$dowlknum=="na"] <- NA
iw$dowlknum[iw$dowlknum=="none"] <- NA
iw$dowlknum[iw$dowlknum=="NONE"] <- NA

iw$dowlknum[iw$dowlknum=="none, part of Winnibigoshish"] <- "11-0147"

# There are some dowlknum that use hyphens, but not according to the convention above
# These all look good: CC-LLLL
unique(iw$dowlknum[grepl("-", iw$dowlknum) & nchar(iw$dowlknum)==7])
# None here:
unique(iw$dowlknum[grepl("-", iw$dowlknum) & nchar(iw$dowlknum)==8])
# Here's a mistake where there isn't a second hyphen
unique(iw$dowlknum[grepl("-", iw$dowlknum) & nchar(iw$dowlknum)==9])
# fix second hyphen
iw$dowlknum[iw$dowlknum=="18-012601"] <- "18-0126-01"
# These all look good; CC-LLLL-BB
unique(iw$dowlknum[grepl("-", iw$dowlknum) & nchar(iw$dowlknum)==10])

iw$parentdow <- substr(iw$dowlknum, 1, 7)

## Walleye lakes----
### UPDATE FILE PATH WHERE WAE DATA IS STORED ----
wae_lakes <- read_csv("R and stan code/Data/wae_NR_class.csv")


## Getting wae_lakes and iw to work together----

unique(iw$dowlknum)
unique(iw$parentdow)

# Looks like a mostly clean list of DOWs with the leading zero problem + a couple combined entries
unique(wae_lakes$DOW)
wae_lakes$DOW[substr(wae_lakes$DOW, 1, 1)=="0"]

# Breaking down what kinds of entries are in here
unique(wae_lakes$DOW[nchar(wae_lakes$DOW)<=6])
unique(wae_lakes$DOW[nchar(wae_lakes$DOW)==7])# these all look like counties 01 - 09 without the leading zero
print(unique(wae_lakes$DOW[nchar(wae_lakes$DOW)==8]), max=1500)# these look like clean, 8 digit DOWs
unique(wae_lakes$DOW[nchar(wae_lakes$DOW)>=9])# these should be fairly easy to account for

iw%>%filter(dowlknum%in%"18-0269" & species=="zebra mussel")# infested
iw%>%filter(dowlknum%in%"18-0268" & species=="zebra mussel")# infested
iw%>%filter(dowlknum%in%"18-0311" & species=="zebra mussel")# infested
iw%>%filter(dowlknum%in%"18-0270" & species=="zebra mussel")# not known to be infested

# Pulling hyphens for infested waters, adding "00" for entries with just parent DOWs, 
# then dropping leading zeros, should do the trick to match "iw" to "wae_lakes"
iw_wae_fixed <- iw %>%
  mutate(DOW=str_remove_all(dowlknum, "-")) %>%
  mutate(DOW=ifelse(nchar(DOW)==6, paste(DOW, "00", sep=""), DOW)) %>%
  mutate(DOW=ifelse(substr(DOW, 1, 1)=="0", substr(DOW, 2, 8), DOW)) %>%
  filter(!(is.na(year) | is.na(species) | is.na(dowlknum)))

unique(iw_wae_fixed$DOW[nchar(iw_wae_fixed$DOW)<=6])
unique(iw_wae_fixed$DOW[nchar(iw_wae_fixed$DOW)==7])# these all look like counties 01 - 09 without the leading zero
unique(iw_wae_fixed$DOW[nchar(iw_wae_fixed$DOW)==8])# these look like clean, 8 digit DOWs
unique(iw_wae_fixed$DOW[nchar(iw_wae_fixed$DOW)>=9])# these should be fairly easy to account for

# Identifying which infested waters lakes have walleye
iw_wae_fixed$wae <- ifelse(iw_wae_fixed$DOW%in%wae_lakes$DOW, 1, 0)

# The combined dow entries above will have wae=0 here, so need to change that manually
iw_wae_fixed$wae[iw_wae_fixed$parentdow=="18-0269"] <- 1
iw_wae_fixed$wae[iw_wae_fixed$parentdow=="18-0268"] <- 1
iw_wae_fixed$wae[iw_wae_fixed$parentdow=="18-0311"] <- 1

# Filtering down to confirmed zebra mussel infestations, then grouping by parentdow
# and summarizing to the earliest designation for a parent dow and wae presence
iw_wae_zm_parentdow <- iw_wae_fixed %>%
  filter(species=="zebra mussel" & connected=="confirmed") %>%
  group_by(parentdow) %>%
  summarize(year=min(year), 
            wae=max(wae)) %>%
  ungroup()

# Grouping by year, estimating number of infestations in that year, then 
# taking the cumulative sum of infestations.
iw_wae_zm_parentdow_yrsum <- iw_wae_zm_parentdow %>%
  group_by(year) %>%
  summarize(n=n()) %>%
  ungroup() %>%
  mutate(cumsum_n=cumsum(n))

# Doing the same, but for walleye lakes
iw_wae_zm_parentdow_yrsum_wae <- iw_wae_zm_parentdow %>%
  filter(wae==1) %>%
  group_by(year) %>%
  summarize(n=n()) %>%
  ungroup() %>%
  mutate(cumsum_n=cumsum(n))

# Summaries and data viz ----

# There are 299 lakes (unique 6-digit DOWs) with ZM and
# 231 of those lakes are walleye lakes.
iw_wae_zm_parentdow_yrsum %>% filter(year==2022) #299
iw_wae_zm_parentdow_yrsum_wae %>% filter(year==2022) #231

# That's 77% (77.26) of all ZM lakes
231/299

# Up until 2019 
iw_wae_zm_parentdow_yrsum %>% filter(year==2019) #229
iw_wae_zm_parentdow_yrsum_wae %>% filter(year==2019) #176

#77 (76.86) % of lakes invaded with ZM as of 2019 are walleye lakes
176/229

# There are 1435 known walleye lakes in the state.
# There are 11,842 lakes > 10 acres in the state. https://www.dnr.state.mn.us/faq/mnfacts/water.html

# Statewide ZM known infestation rate = 2.5%, 11,842 lakes in MN that are 10+ acres
299/11842

# ZM known infestation rate in known walleye lakes = 16% 
231/1435

# ZM known infestation rate in known walleye lakes through 2019 = 12.26%
176/1435

# ZM known infestation rate in lakes not known to have walleye = 0.7%
68/10407

# ZM known infestation rate in lakes not known to have walleye = 0.5% through 2019
(229-176)/(11842-1435)

# ZM are nearly 25 times more likely to be found in lakes known to contain walleye
(231/1435)/(68/10407)

# ZM are 24 times more likely to be found in lakes known to contain walleye through 2019
(176/1435)/(53/10407)

# number of ZM invaded lakes through 2019
ZM_WAE_2019 <- iw_wae_zm_parentdow_yrsum_wae %>%
  filter()
```

