---
title: "gn_WAE_ZM_Model1"
author: "Holly Kundel"
date: "11/15/2021"
output: html_document
---

Load in Required Packages

```{r, warings = F, messages = F}
#remove.packages(c("StanHeaders", "rstan"))
#install.packages("rstan", repos = c('https://stan-dev.r-universe.dev', getOption("repos")))
#if (file.exists(".RData")) file.remove(".RData")
#This one works best?
#install.packages("rstan", repos = "https://cloud.r-project.org/", dependencies = TRUE)
library("rstan") # observe startup messages
library(readr)
library(dplyr)
library(tidyr)
library(lubridate)
library(stringr)
library(arm)
library(doBy)
library(sf)
library(spData)
library(RANN)
library(ggplot2)
library(MASS)
library(bayesplot)
library(brms)
library(rstan)
library(bayesplot)
library(lme4)
library(gridExtra)
library(ggthemes)
library(coda)
library(forcats)
options(scipen = 999)
```

Load in Data

```{r, message=FALSE}
dat <- read_csv("./cleaned_data/year_3_fish_lake_age_alk.csv") %>% 
  mutate(survey_id = row_number()) %>% drop_na()

unique(dat$lake.id)


with(dat, hist(age3_count), breaks = 1000)
nrow(dat[dat$age3_count==0,])/nrow(dat)
```

# Nearest Neighbor Code

-   Assigning uninvaded lakes a "pseudo-invasion year" for BACI set up

Data setup
Create Spatial sf Object

```{r}
# Create spatial sf object (4326 = WGS84)
dat.geo <- st_as_sf(dat, coords = c("x", "y"), crs = 4326)
```

Separate Invaded Lakes and Uninvaded Lakes

-   Using column `ZM_LAKE` for this
-   Only considered invaded if at least one gn survey was done the year of invasion or later
-   
-   "0" indicates a lake is uninvaded (based on above criteria)
-   "1" indicates a lake is invaded (based on above criteria)

```{r}
dat.geo.uninvaded <- dat.geo %>% 
  filter(year_confirmed == 0)
dim(dat.geo.uninvaded) 

dat.geo.invaded <- dat.geo %>% 
  filter(year_confirmed > 0)
dim(dat.geo.invaded) 
```

Get coordinate matrices

```{r}
uninvaded_coords <- do.call(rbind, st_geometry(dat.geo.uninvaded))
invaded_coords <- do.call(rbind, st_geometry(dat.geo.invaded))
invaded_coords <- cbind(invaded_coords, 1:nrow(invaded_coords)) # Index for identifying nearest lake

```

Find nearest neighbor

```{r}
closest <- nn2(invaded_coords[,1:2], uninvaded_coords,  k = 1, searchtype = "standard")
#str(closest)
#head(closest)
closest <- sapply(closest, cbind) 
```

Combine uninvaded data with nearest neighbor assignments

```{r}
dat.geo.uninvaded <- cbind(dat.geo.uninvaded, closest)
dim(dat.geo.uninvaded) 
```

Give invaded lakes an identifier and select columns

```{r}
dat.geo.invaded$nn.idx <- 1:nrow(dat.geo.invaded)
dat.geo.invaded <- dat.geo.invaded %>% 
  dplyr::select(year_confirmed, nn.idx)
head(dat.geo.invaded)

#str(dat.geo.uninvaded)
#str(dat.geo.invaded)
```

Merge uninvaded with invaded `YEAR_INFESTED` for use as "after" year in BACI (Before After Control Impact)

```{r}
dat.geo.invaded_merge <- st_set_geometry(dat.geo.invaded, NULL)

dat.geo.uninvaded.final <- dat.geo.uninvaded %>%
    left_join(dat.geo.invaded_merge, by = c("nn.idx"))
str(dat.geo.uninvaded.final)
summary(dat.geo.uninvaded.final)
# Bring back into full dataset
# Convert to dataframe
dat.geo.uninvaded.final.df <- st_set_geometry(dat.geo.uninvaded.final, NULL)
class(dat.geo.uninvaded.final.df)
str(dat.geo.uninvaded.final.df)

# Join dat.geo.uninvaded.final.df - want YEAR_INFESTED.y - with original data, dat
# Grab columns of interest
dat.geo.uninvaded.final.df <- dat.geo.uninvaded.final.df %>% 
  dplyr::select(survey_id, year_confirmed.y)
dim(dat.geo.uninvaded.final.df)
head(dat.geo.uninvaded.final.df)
dim(dat)

# Final merge
final.dat3 <- dat %>%
    left_join(dat.geo.uninvaded.final.df, by = c("survey_id"))
dim(final.dat3)
head(final.dat3)
```


# Prepare to run model
```{r}
dat <- final.dat3 %>%
  mutate(ZM_LAKE = ifelse(year_confirmed == 0, 0, 1),
         log_acres = log(acres),
         fry.pa = FRY/acres,
         fgl.pa = FGL/acres, 
         offsets = log(total.effort.1))
```

Create BACI design variables

```{r}
# "Treatment" is lake-level invasion status, ZM (1 = ZM; 0 otherwise)
# Before/After time variable ("post" = 1 if sampled after invasion; 0 if before)
# Below, "year" is year of sampling and time is year infested
dat <- dat %>% 
  mutate(time = ifelse(year_confirmed == 0, year_confirmed.y, year_confirmed),
         post = ifelse(birth.year >= time, 1, 0),
         treatment = ZM_LAKE # USING ZM_LAKE_HERE
         )%>%
  mutate(INVASION_STATUS_NEW = case_when(ZM_LAKE == 0 ~ "Uninvaded",
                                         ZM_LAKE == 1 & post == 0 ~ "PRE_ZM",
                                         ZM_LAKE == 1 & post == 1 ~ "POST_ZM"))


#check data
dat %>%
  group_by(ZM_LAKE) %>%  
  count(post, ZM_LAKE) %>%
  ungroup() %>% 
  mutate(prop = prop.table(n))

###new addition of data filtering#######
#ensure at least one year pre and post
test = dat %>% 
  group_by(lake.id, post, ZM_LAKE) %>% 
  summarise(count=n()) %>% 
  filter(count>0) %>% ungroup() %>% 
  group_by(lake.id, ZM_LAKE) %>% 
  summarise(pre.post.test=n()) %>% 
  filter(pre.post.test==2)


dat <- dat %>% filter(lake.id %in% test$lake.id)
```

Lake-level predictors

```{r}
# Prepare lake-level predictors
lake_area <- summaryBy(log_acres ~ lake.id, data=dat, FUN=mean)
gdd <- summaryBy(median_gdd_5 ~ lake.id, data=dat, FUN=mean)


# Transform as needed and scale
lake_area2 <- as.numeric(scale(log(lake_area$log_acres.mean)))
gdd2 <- as.numeric(scale(gdd$median_gdd_5.mean))
```

Observation-level predictors

```{r}
# Observation-level predictors, transform as needed and scale
sum <- (dat$gdd_wtr_5c_year0 + dat$gdd_wtr_5c_year1 + dat$gdd_wtr_5c_year2) - (3*dat$median_gdd_5)
gdd_sum <- as.numeric(scale(sum))
fry.pa <- as.numeric(scale(dat$fry.pa))
fgl.pa <- as.numeric(scale(dat$fgl.pa))

hist(fry.pa); hist(fgl.pa)

post <- dat$post
x.treatment <- dat$treatment
```

Quick test with glmer
```{r, eval = F}
m0 <-glmer(age3_count~ 1 + (1 | lake.id), data = dat, family = poisson(link = "log"), nAGQ = 100)
#summary(m0)
# # 
m1 <- glmer(age3_count ~ post + treatment + post*treatment + (1|lake.id), nAGQ = 100,
             data=dat, family = poisson(link="log"))
#summary(m1)
performance::check_zeroinflation(m1)
```

Export data with correct number of observations and scaled variables

```{r}
dat$gdd_sum_S <- gdd_sum
dat$fry.pa_S <- fry.pa
dat$fgl.pa_S <- fgl.pa

#write.csv(dat, "./cleaned_data/gn_model_data.csv", row.names = F)
```



# STAN Model
```{r, eval = F}
data <- list(y = dat$age3_count, lake = dat$lake.id, N = dim(dat)[1], 
             n_lake = n_lake, offsets=dat$offsets, 
             year = dat$year.id, n_year = n_year, n_post = n_post, n_trt = n_trt,
             post = post, trt = x.treatment, 
             fry_pa = fry.pa, fgl_pa = fgl.pa,
             survey_gdd_sum = gdd_sum, gdd = gdd2, 
             lake_area = lake_area2 
             ) 

options(scipen = 999)
out <- stan(file='./pois_model_zip3_GN.stan', data = data,
            iter = 50000, warmup = 20000, chains=4, thin=2, cores=4, refresh = 1)
print(out)


# Save all the stan output
#
saveRDS(out, file="./models/model_out_zip_GN_nosecchi_3gddanomalyCR.rds")
out <- readRDS(file="./models/model_out_zip_GN_nosecchi_3gddanomalyCR.rds")
summary(out)

mcmc_trace(out, pars = c("sigma_year", "sigma_lake", "theta", 
                         "b_0", "b_post", "b_trt",  "b_post_trt", "b_fry", "b_fry", "b_fgl", "b_lake_area"))
```



```{r}
to_keep <- c("anomaly_gdd_5_s","anomaly_gdd_5", "fry", "fingerling", "median_gdd_5_s", "median_gdd_5", "lake_area_s", "log_acres")

dat_standardized <- dat %>% 
  mutate(anomaly_gdd_5 = as.numeric((gdd_wtr_5c_year0 + gdd_wtr_5c_year1 + gdd_wtr_5c_year2) - 3*median_gdd_5),
         anomaly_gdd_5_s = as.numeric(scale((gdd_wtr_5c_year0 + gdd_wtr_5c_year1 + gdd_wtr_5c_year2) - 3*median_gdd_5)),
         lake_area_s = as.numeric(scale(log_acres)),
         median_gdd_5_s = as.numeric(scale(median_gdd_5)),
         fry = as.factor(ifelse(fry.pa == 0, 0, 1)),
         fingerling = as.factor(ifelse(fgl.pa == 0, 0, 1)),
         year = as.factor(year)) %>%
  dplyr::select(lake.id, year, age3_count, offsets, total.effort.1, post, treatment, x, y, matches(to_keep))

write.csv(dat_standardized, "./cleaned_data/gn_for_model.csv", row.names = F)
dat_standardized <- read.csv("./cleaned_data/gn_for_model.csv")

temp <- dat_standardized %>% group_by(lake.id) %>% count()
median(temp$n)

length(unique(temp$lake.id))
min(dat_standardized$year); max(dat_standardized$year)
```


```{r}
priors = c(
    prior(normal(0, 5), class = "Intercept"),
    prior(normal(0, 5), class = "b")
  )

### Zero inflated poisson with covariates
zip_gn_model <- 
  brm(age3_count ~ offset(offsets) + 
      post + treatment + post:treatment + fry + fingerling + 
      anomaly_gdd_5_s + median_gdd_5_s + 
      lake_area_s + median_gdd_5_s +
      (1 | year) +              
      (1 | lake.id),
    family = zero_inflated_poisson(), data = dat_standardized, prior = priors,
    control = list(max_treedepth = 25),
    chains = 4, iter = 10000, warmup = 2500, thin = 2, seed = 1024, save_pars=save_pars(all=TRUE))
#summary(zip_gn_model)
#plot(zip_gn_model)
saveRDS(zip_gn_model, "./models/zip_gn_model_brms.rds")
zip_gn_model <- readRDS("./models/zip_gn_model_brms.rds")


### Negative Binomial with covariates
nb_gn_model <- 
  brm(age3_count ~ offset(offsets) + 
      post + treatment + post:treatment + fry + fingerling + 
        anomaly_gdd_5_s + median_gdd_5_s + 
      lake_area_s + median_gdd_5_s +
      (1 | year) +              
      (1 | lake.id), 
    family = negbinomial(), data = dat_standardized, prior = priors,
    chains = 4, iter = 10000, warmup = 2500, thin = 2, seed = 1024, save_pars=save_pars(all=TRUE))
#plot(nb_gn_model)
#summary(zip_gn_model)
saveRDS(nb_gn_model, "./models/nb_gn_model_brms.rds")
#nb_gn_model <- readRDS("./models/nb_gn_model_brms.rds")



### Zero inflated poisson NO covariates
zip_gn_model_nocovs <- 
  brm(age3_count ~ offset(offsets) + 
      post + treatment + post:treatment +
      (1 | year) +              
      (1 | lake.id), 
    family = zero_inflated_poisson(), data = dat_standardized, prior = priors,
    control = list(max_treedepth = 25),
    chains = 4, iter = 10000, warmup = 2500, thin = 2, seed = 1024, save_pars=save_pars(all=TRUE))
saveRDS(zip_gn_model_nocovs, "./models/zip_gn_model_nocovs_brms.rds")
rm(zip_gn_model_nocovs)

### Negative Binomial NO covariates
nb_gn_model_nocovs <- 
  brm(age3_count ~ offset(offsets) + 
      post + treatment + post:treatment +
      (1 | year) +              
      (1 | lake.id), 
    family = negbinomial(), data = dat_standardized, prior = priors,
    chains = 4, iter = 10000, warmup = 2500, thin = 2, seed = 1024, save_pars=save_pars(all=TRUE))
saveRDS(nb_gn_model_nocovs, "./models/nb_gn_model_nocovs_brms.rds")
#nb_gn_model_nocovs <- readRDS("./models/nb_gn_model_nocovs_brms.rds")
#summary(nb_gn_model_nocovs)
rm(nb_gn_model_nocovs)

### Zero inflated Negative Binomial with covariates
### USED IN PAPER
zipnb_gn_model <- 
  brm(age3_count ~ offset(offsets) + 
      post + treatment + post:treatment + fry + fingerling + 
      anomaly_gdd_5_s + median_gdd_5_s + 
      lake_area_s + 
      (1 | year) +              
      (1 | lake.id), 
    family = zero_inflated_negbinomial(), data = dat_standardized, prior = priors,
    chains = 4, iter = 10000, warmup = 2500, thin = 2, seed = 1024, save_pars=save_pars(all=TRUE))
#plot(zipnb_gn_model)
#summary(zipnb_gn_model)
saveRDS(zipnb_gn_model, "./models/zipnb_gn_model_brms.rds")
#zipnb_gn_model <- readRDS("./models/zipnb_gn_model_brms.rds")
```


# Model selection, and other fun tidbits
```{r}
# These take quite a bit to run and dont work (high pareto K value)
output_nb <- loo(nb_gn_model, moment_match = TRUE)
output_zipnb <- loo(zipnb_gn_model, moment_match = TRUE)
output_zip <- loo(zip_gn_model, moment_match = TRUE)

loo_compare(output_nb, output_zipnb, output_nb_nocovs)

options(future.globals.maxSize= 891289600)
start <- Sys.time()
kfoldzip <- kfold(zip_gn_model, k = 10) # elpd_kfold = -7022.6	SE = 411.7	 	
#kfoldnb <- kfold(nb_gn_model, k = 10) # elpd_kfold = -3315.9		SE = 58.6	
#kfoldzionb <- kfold(zipnb_gn_model, k = 10) # elpd_kfold = -3311.7		SE = 	58.4
#print(kfoldzip); print(kfoldnb); print(kfoldzionb)
kfold_results <- list(zip = kfoldzip, nb = kfoldnb, zinb = kfoldzionb)
#saveRDS(kfold_results, "./models/kfold_results_gn.rds")
end <- Sys.time()
total <- end - start

kfold_results$zip
kfold_results$nb
kfold_results$zinb

summary(zip_gn_model)         
#post:treatment = -0.24
summary(nb_gn_model)          
#post:treatment = -0.18
summary(zip_gn_model_nocovs)
#post:treatment =
summary(nb_gn_model_nocovs)
#post:treatment = -0.56
summary(zipnb_gn_model)
#post:treatment = -0.18

bayes_R2(nb_gn_model)
pp_check(nb_gn_model, ndraws = 100)
pp_check(nb_gn_model, type='stat', stat='mean')
pp_check(nb_gn_model, type='error_scatter_avg')

bayes_R2(zipnb_gn_model) # 0.535
pp_check(zipnb_gn_model, ndraws = 100)
pp_check(zipnb_gn_model, type='stat', stat='mean')
pp_check(zipnb_gn_model, type='error_scatter_avg')
```

